### 2023-03-17 paper/main.old.tex
use the springer template

\documentclass[fleqn,10pt]{olplainarticle}

\title{Predicting Audio Features with Last.fm Tags}

\author[1]{Jaime Ramírez Castillo}
\author[1]{M. Julia Flores}
\affil[1]{Departamento de Sistemas Informáticos, Universidad de Castilla - La Mancha, Spain}

\keywords{Music information retrieval, Artificial intelligence}

\begin{abstract}
In this paper, we discuss a number of experiments to analyze the suitability of music label representations to predict certain audio features, such as danceability, loudness, or acousticness...
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}

Music information retrieval (MIR) is a field that focuses on the analysis, processing and knowledge discovery of information stored in music pieces \cite{ramirez2020machine}.

Many of the challenges in this field use audio features, such as valence and energy, to predict other music-related aspects.
For example, some studies have been using these features to predict the probably of a track being a hit [REF].

To make these predictions, researchers have been using machine learning models, specially over the past few years.
Previously, researchers commonly used hand-crafted audio computational models to perform these tasks.

But what if we tried to use these audio features as the predicted values? We would need a set of predictor variables to
acurrately predict the value of audio features.
This premise is the core concept of this article.

The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.

In the remainder of the article, we explain the data gathering and preparation process, as well as the data input formats and varios models.
We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.

\section*{State of the Art}

In recent years, researchers have used Last.fm tags in classification and regression tasks.
Several studies have used Last.fm to predict music sentiment or mood.

% https://core.ac.uk/download/pdf/234915638.pdf

% https://sciendo.com/pdf/10.2478/ausi-2018-0009

% https://archives.ismir.net/ismir2009/paper/000018.pdf

Additionally, the Spotify audio features...

In general, these studies confirm the posibility of extracting knowledge from Last.fm tags.
To the best of our knowledge, no studies have addressed the problem of audio features regression, based solely on Last.fm tags.

\textcolor{red}{To be Continued...}

\section*{Generating a Dataset}

Before conducting experiments on predicting audio features from tags, we constructed a dataset, by gathering the data from the Last.fm and Spotify APIs.

\subsection*{A Single-user Dataset}

Similar to other intelligent systems, recommender systems
must be trained, by using user preference data, to produce
adequate recommendations. For our recommendation framework,
we have leveraged the knowledge discovery potential
of large historical listening logs, gathered from Last.fm.

To characterize the preferences and context of the user, we
have chosen to start with a simple scenario, where just data
from a single user is available. By training our system with
data from a single user, we also want to begin a discussion,
given the following question: Is it possible to train 
recommender systems, and in particular, user-centric systems, by
using a single-user dataset?

To the best of our knowledge, research on user-centric rec-
ommender systems has concentrated its efforts on explain-
able AI Wang et al. [2019], and also user-centered evaluation
of these systems Knijnenburg et al. [2012]. We also argue
that recommender systems that exploit the preferences of a
single user, or a reduced number of users, might as well be
considered as user-centric models.

\subsection*{Last.fm}

Last.fm is a online music service for uses to keep track of their music listenting habits.
Last.fm can also be considered as an community where users tag artists, albums, and tracks, according the the own perception of the user.

For nearly two decades, users have been contributing to Last.fm by tagging tracks with arbitrary text labels.
These tags do not necessarily have to be single-worded.
Users often use short sentences to define a song, such as `I like this track`, or `on the beach`.

Community-contributed tags from the Last.fm API.
These tags are text labels that Last.fm users assign
to artists, albums, or tracks. Users apply these tags to
categorize music from their own perspective, which
means that tags do not fit into any structured ontology
or data model. Tags can refer to aspects such as genre,
emotion, or user context.

\subsubsection*{Last.fm Tags}
Last.fm uses the term \emph{scrobble} to refer to a single track playback,
in a particular moment. We have queried the Last.fm
API to download the user{'}s scrobble logs, reported from 2007 to 2022.
For each scrobble, we have gathered the following information:

\begin{itemize}
     \item Track playback timestamp.
     \item Track MusicBrainz Identifier (MBID), if exists.
     \item Track name
     \item Artist name
     \item Track tags. If the track does not have any tags assigned,
      then artist tags have been used.
\end{itemize}

For each tag assigned to a track, or an artist, Last.fm includes
a count property to indicate the popularity of the given tag for the track.
Last.fm normalizes this value in the 0-100 range, so the most popular tag for a track can have a
count value of 100.

Users normally listens to their favorite tracks many times,
so the amount of individual tracks listened is much smaller
than the number of track plays. In this case, the amount of
individual tracks listened is about 20,000.

The format is as follows:

\begin{verbatim}
{
      "artist - name": {
            "eletronica": 100,
            "rock": 80,
            "pop": 45,
            "jazz": 0,
            "nu-jazz": 0,
            "country": 0,
            "soul": 0,
      }
}
\end{verbatim}

\subsection*{Spotify}

After gathering Last.fm data and identifying the unique
tracks that represent the user music collection, we have
collected Spotify audio features. For each of these individ-
ual tracks, we have downloaded the Spotify audio features
specific to the given track.

The Spotify audio features are numerical values that repre-
sent high-level audio information computed from a specific
track. These values characterize a track, musically speaking,
by measuring relevant musical aspects.

The features provided by the Spotify API are acousticness,
danceability, duration\textunderscore ms, energy, instrumentalness, key,
liveness, loudness, mode, speechiness, tempo, and valence.
Table 1 describes these features. The reader can find further
details about each feature in the Spotify API documentation
4.

A small portion of the tracks do not have features available in
Spotify, so they have been filtered out from our experiments.

After filtering songs that miss Last.fm tags or Spotify audio features,
our dataset contains 14009 samples.

\emph{Track audio features from the Spotify API}. These are
attributes computed from the audio themselves. They
are a way to describe music by using numerical values.
For example, a danceability attribute of 0.95 means
that a particular song is highly suitable for dancing.

\subsection*{Last.fm Tags Representations for Training}

\subsubsection*{Tabular}
Each tag is a column and each cell contains the popularity value of a tag for a track.
A cell is 0 if a tag is missing for a track.

The number of columns is limited to the top-K tags.

\subsubsection*{Tabular Tokens}

Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
To tokenize tags, we have used the GTP2 tokenizer.
Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
To \emph{stringify} the tags, we have concatenated tags with multiple strategies:

\begin{itemize}
      \item By including tag popularity: `rock 2, pop 1`.
      \item By repeating tags based on popularity: `rock rock, pop`.
      \item By ordering by popularity: `rock, pop`.
\end{itemize}

\subsection*{Training Data By Track}

When generating training data by track, the tabular formats present sparsity problems.

For tabular representations, we need to defined a fixed set of columns as tags.
For most of tracks, most columns are `0`.

The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
(e.g., m * n for an m * n matrix) is called the sparsity of the matrix

\section*{Experiments}

\textcolor{red}{TODO}

\subsection*{Models}

\textcolor{red}{TODO}

\section*{Conclusions}

\textcolor{red}{TODO}

\section*{Acknowledgments}

\textcolor{red}{TODO}

\bibliography{sample}

\end{document}

### 2023-03-17 paper/main.tex
use the springer template

--- 2023-03-17 paper/main.old.tex
@@
-\documentclass[fleqn,10pt]{olplainarticle}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\title{Predicting Audio Features with Last.fm Tags}
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
 
-\author[1]{Jaime Ramírez Castillo}
-\author[1]{M. Julia Flores}
-\affil[1]{Departamento de Sistemas Informáticos, Universidad de Castilla - La Mancha, Spain}
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
+
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
+
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
+%%\documentclass[default]{sn-jnl}% Default
+%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published
+%%%%  by Springer Nature. The guidance has been prepared in partnership with
+%%%%  production teams to conform to Springer Nature technical requirements.
+%%%%  Editorial and presentation requirements differ among journal portfolios and
+%%%%  research disciplines. You may find sections in this template are irrelevant
+%%%%  to your work and are empowered to omit any such section if allowed by the
+%%%%  journal you intend to submit to. The submission guidelines and policies
+%%%%  of the journal take precedence. A detailed User Manual is available in the
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+\jyear{2023}%
+
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}%
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
+
+% JAIME: this is to hide font size warnings
+\usepackage{anyfontsize}
+
+\begin{document}
+
+\title[Predicting Audio Features with Last.fm Tags]{Predicting Audio Features with Last.fm Tags}
+
+%%=============================================================%%
+%% Prefix	-> \pfx{Dr}
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% NatureName	-> \tanm{Poet Laureate} -> Title after name
+%% Degrees	-> \dgr{MSc, PhD}
+%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
+%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\author*[1]{\fnm{Jaime} \sur{Ramírez Castillo}}\email{Jaime.Ramirez@alu.uclm.es}
+
+\author[1]{\fnm{M. Julia} \sur{Flores}}\email{Julia.Flores@uclm.es}
+\equalcont{These authors contributed equally to this work.}
+
+% Departamento de Sistemas Informáticos, Universidad de Castilla - La Mancha, Spai
+
+\affil[1]{\orgdiv{Departamento de Sistemas Informáticos}, \orgname{Universidad de Castilla-La Mancha}, \orgaddress{\street{Campus universitario s/n}, \city{Albacete}, \postcode{02071}, \country{Spain}}}
+
+%%==================================%%
+%% sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{
+    In this paper, we discuss a number of experiments to analyze the
+    suitability of music label representations to predict certain audio features,
+    such as danceability, loudness, or acousticness...
+}
+
+%%================================%%
+%% Sample for structured abstract %%
+%%================================%%
+
+% \abstract{\textbf{Purpose:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+%
+% \textbf{Methods:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+%
+% \textbf{Results:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+%
+% \textbf{Conclusion:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}
 
@@
 
-\begin{abstract}
-In this paper, we discuss a number of experiments to analyze the suitability of music label representations to predict certain audio features, such as danceability, loudness, or acousticness...
-\end{abstract}
+%%\pacs[JEL Classification]{D8, H51}
 
-\begin{document}
+%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}
 
-\flushbottom
 \maketitle
-\thispagestyle{empty}
 
-\section*{Introduction}
+\section{Introduction}\label{sec1}
 
-Music information retrieval (MIR) is a field that focuses on the analysis, processing and knowledge discovery of information stored in music pieces \cite{ramirez2020machine}.
+Music information retrieval (MIR) is a field that focuses on the analysis,
+processing and knowledge discovery of latent information in music pieces \cite{ramirez2020machine}.
 
@@
 
-\section*{State of the Art}
+\section{State of the Art}
 
@@
 
-\section*{Generating a Dataset}
+\section{Generating a Dataset}
 
@@
 
-\subsection*{A Single-user Dataset}
+\subsection{A Single-user Dataset}
 
@@
 
-\subsection*{Last.fm}
+\subsection{Last.fm}
 
@@
 
-\subsubsection*{Last.fm Tags}
+\subsubsection{Last.fm Tags}
 Last.fm uses the term \emph{scrobble} to refer to a single track playback,
@@
 
-\subsection*{Spotify}
+\subsection{Spotify}
 
@@
 
-\subsection*{Last.fm Tags Representations for Training}
+\subsection{Last.fm Tags Representations for Training}
 
-\subsubsection*{Tabular}
+\subsubsection{Tabular}
 Each tag is a column and each cell contains the popularity value of a tag for a track.
@@
 
-\subsubsection*{Tabular Tokens}
+\subsubsection{Tabular Tokens}
 
@@
 
-\subsection*{Training Data By Track}
+\subsection{Training Data By Track}
 
@@
 
-\section*{Experiments}
+\section{Experiments}
 
@@
 
-\subsection*{Models}
+\subsection{Models}
 
@@
 
-\section*{Conclusions}
+\section{Conclusions}
 
@@
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 
@@
 
-\bibliography{sample}
+%%===========================================================================================%%
+%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
+%% system, please include the references within the manuscript file itself. You may do this  %%
+%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
+%% file, and delete the associated \verb+\bibliography+ commands.                            %%
+%%===========================================================================================%%
+
+\bibliography{sn-bibliography}% common bib file
+%% if required, the content of .bbl file can be included here once bbl is generated
+%%\input sn-article.bbl
+
+%% Default %%
+%%\input sn-sample-bib.tex%
 

### 2023-03-19 paper/main.tex
add input data tables

@@
 %%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
+%% Please do not use \input{\dots} to include other tex files.       %%
 %% Submit your LaTeX manuscript as one .tex document.              %%
@@
 
-% Departamento de Sistemas Informáticos, Universidad de Castilla - La Mancha, Spai
-
 \affil[1]{\orgdiv{Departamento de Sistemas Informáticos}, \orgname{Universidad de Castilla-La Mancha}, \orgaddress{\street{Campus universitario s/n}, \city{Albacete}, \postcode{02071}, \country{Spain}}}
@@
     suitability of music label representations to predict certain audio features,
-    such as danceability, loudness, or acousticness...
+    such as danceability, loudness, or acousticness \dots 
 }
@@
 
-Additionally, the Spotify audio features...
+Additionally, the Spotify audio features \dots
 
@@
 
-\textcolor{red}{To be Continued...}
+\textcolor{red}{To be Continued \dots }
 
@@
 
+\subsection{Formatting Input Data for Predicting From Last.fm Tags}
+
+The input data passed to the XGBoost regressor is formatted in tabular format, as follows:
+
+\begin{itemize}
+      \item Given that $Tags_{k}$ is the set of most $k$ frequent Last.fm tags in the user listening history
+            and, where each $tag \in Tags_{k}$.
+      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
+      \item For each $track$:
+      \begin{itemize}
+            \item $X_{track},_{tag}$ is the strengh of $tag$ for $track$. This value is in the $0-100$ range.
+            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
+      \end{itemize}
+\end{itemize}
+
+An example of this data format is provided in table \ref{tabular_tags_format}.
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Tabular data format for Last.fm tags in XGBoost and Bayesian regressors}\label{tabular_tags_format}%
+      \begin{tabular}{@{}lllllll@{}}
+      \toprule
+      Track                         & $X_{electronic}$ & $X_{ambient}$ & $X_{\dots}$ & $y_{energy}$ & $y_{valence}$ & $y_{\dots}$ \\
+      \midrule
+      Massive Attack - Blue Lines   & 62               & 6             &  \dots      & 0.496        & 0.947         & \dots  \\
+      The Beta Band - Squares       & 40               & 3             &  \dots      & 0.446        & 0.507         & \dots  \\
+      \dots                         & \dots            & \dots         &  \dots      & \dots        & \dots         & \dots  \\
+      \botrule
+      \end{tabular}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+\subsection{Formatting Input Data for Predicting From Tokens}
+
+In this particular case, the $X$ values of the tabular input data are tokens.
+These tokens are obtained from passing the a string of concatenated Last.fm tags through a tokenizer.
+The formal definition of this data format is as follows:
+
+\begin{itemize}
+      \item Given that $X_l$ is the token vocabulary, where $l$ is the maximum vocabulary length.
+      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
+      \item For each $track$:
+      \begin{itemize}
+            \item $X_{track},_{n}$ is token found at position $n$, after tokenizing the tags string.
+            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
+      \end{itemize}
+\end{itemize}
+
+An example of this data format is provided in table \ref{tabular_token_format}.
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Tabular data format for tokens in XGBoost and Bayesian regressors}\label{tabular_token_format}%
+      \begin{tabular}{@{}llllllll@{}}
+      \toprule
+      Track                         & $X_{0}$ & $X_{1}$ & $X_{2}$ & $X_{\dots}$ & $y_{energy}$ & $y_{valence}$ & $y_{\dots}$ \\
+      \midrule
+      Massive Attack - Blue Lines   & 101     & 5099    & 6154    &  \dots      & 0.496        & 0.947         &  \dots  \\
+      The Beta Band - Squares       & 101     & 4522    & 2600    &  \dots      & 0.446        & 0.507         &  \dots  \\
+      \dots                         & \dots   & \dots   & \dots   &  \dots      & \dots        & \dots         &  \dots  \\
+      \botrule
+      \end{tabular}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+\subsection{Formatting Input Data for Predicting From Text}
+
+When using transformer models, the input data is a string.
+We must represent the Last.fm tags, which are initially in the $(tag name, tag popularity)$ form, to a a string.
+
+After converting to a string, the formal definition of the input data is as follows:
+
+\begin{itemize}
+      \item Given that $X$ is tags represented as text.
+      \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
+      \item For each $track$:
+      \begin{itemize}
+            \item $X_{track},_{n}$ is set of tags for $track$, encoded as a single string.
+            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
+      \end{itemize}
+\end{itemize}
+
+An example of this data format is provided in table \ref{text_format}.
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Text data format for tokens in XGBoost and Bayesian regressors}\label{text_format}%
+      \begin{tabular}{@{}lllll@{}}
+      \toprule
+      Track                         & $X$                                   & $y_{energy}$ & $y_{valence}$ & $y_{\dots}$ \\
+      \midrule
+      Massive Attack - Blue Lines   & "hip hop, chill, bristol, \dots"      & 0.496        & 0.947         & \dots  \\
+      The Beta Band - Squares       & "alternative rock, folk, \dots"       & 0.446        & 0.507         & \dots \\
+      \dots                         & \dots                                 & \dots        & \dots         & \dots  \\
+      \botrule
+      \end{tabular}
+      \end{minipage}
+      \end{center}
+\end{table}
+
 \section{Experiments}
 
-\textcolor{red}{TODO}
+We trained a commonly used machine learning models to predict an audio feature, given the set of tags for a particular track.
 
-\subsection{Models}
+\begin{itemize}
+      \item Boosted tree regressor \cite{xgboost}
+      \item Naive Bayes Regressor \cite{bayesian}
+      \item Fine-tuned GPT-2 model
+\end{itemize}
 
-\textcolor{red}{TODO}
+\subsection{Boosted Tree Regressor}
+
+We configured the boosted tree regressor model with the following training parameters:
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{174pt}
+      \caption{Training parameters for XGBoost regressor}\label{xgboosttable}%
+      \begin{tabular}{@{}llll@{}}
+      \toprule
+      Parameter               & Value \\
+      \midrule
+      objective               & reg:squarederror  \\
+      base score              & 0.5 \\
+      booster                 & gbtree  \\
+      colsample bylevel       & 1 \\
+      colsample bynode        & 1 \\
+      colsample bytree        & 1 \\
+      gamma\footnotemark[1]   & 0 \\
+      learning rate           & 0.300000012 \\
+      max delta step          & 0 \\
+      max depth               & 6 \\
+      min child weight        & 1 \\
+      estimators              & 200  \\
+      n jobs                  & 12  \\
+      num parallel tree       & 1 \\
+      predictor               & auto  \\
+      random state            & 0 \\
+      reg alpha               & 0 \\
+      reg lambda              & 1 \\
+      scale pos weight        & 1 \\
+      subsample               & 2 \\
+      tree method             & auto  \\
+      \botrule
+      \end{tabular}
+      \footnotetext[1]{Minimum loss reduction required to make a further partition on a leaf node of the tree.}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+% \begin{table}[h]
+% \begin{center}
+% \begin{minipage}{\textwidth}
+% \caption{Data format for XGBoost regressor}\label{tab2}
+% \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
+%       \toprule%
+%       & \multicolumn{3}{@{}c@{}}{Tags as columns\footnotemark[1]} & \multicolumn{3}{@{}c@{}}{Tokens as columns\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
+%       Track & rock &  \dots  & $\y_{valence}$ & $\position_{0} &  \dots $ & $\sigma_{valence}$ \\
+%       \midrule
+%       Element 3  & 100 &  \dots  & 0.563 & 780 A & 1166 & $1239\pm100$\\
+%       Element 4  & 75  &  \dots  & 0.434 & 900 A & 1268 & $1092\pm40$\\
+%       \botrule
+%       \end{tabular*}
+%       \footnotetext{Note: This is an example of table footnote.
+This is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
+%       \footnotetext[1]{Example for a first table footnote.}
+%       \footnotetext[2]{Example for a second table footnote.}
+% \end{minipage}
+% \end{center}
+% \end{table}
 
@@
 
-\bibliography{sn-bibliography}% common bib file
+\bibliography{bibliography}% common bib file
 %% if required, the content of .bbl file can be included here once bbl is generated

### 2023-03-19 paper/main.tex
add bayes parameters table

@@
 \usepackage{anyfontsize}
+
+% JAIME: For 1e-10 notation
+\usepackage{siunitx}
 
@@
 
-We configured the boosted tree regressor model with the following training parameters:
+We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.
 
@@
       \begin{minipage}{174pt}
-      \caption{Training parameters for XGBoost regressor}\label{xgboosttable}%
+      \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
       \begin{tabular}{@{}llll@{}}
@@
 
-% \begin{table}[h]
-% \begin{center}
-% \begin{minipage}{\textwidth}
-% \caption{Data format for XGBoost regressor}\label{tab2}
-% \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
-%       \toprule%
-%       & \multicolumn{3}{@{}c@{}}{Tags as columns\footnotemark[1]} & \multicolumn{3}{@{}c@{}}{Tokens as columns\footnotemark[2]} \\\cmidrule{2-4}\cmidrule{5-7}%
-%       Track & rock &  \dots  & $\y_{valence}$ & $\position_{0} &  \dots $ & $\sigma_{valence}$ \\
-%       \midrule
-%       Element 3  & 100 &  \dots  & 0.563 & 780 A & 1166 & $1239\pm100$\\
-%       Element 4  & 75  &  \dots  & 0.434 & 900 A & 1268 & $1092\pm40$\\
-%       \botrule
-%       \end{tabular*}
-%       \footnotetext{Note: This is an example of table footnote.
-This is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote.}
-%       \footnotetext[1]{Example for a first table footnote.}
-%       \footnotetext[2]{Example for a second table footnote.}
-% \end{minipage}
-% \end{center}
-% \end{table}
+\subsection{Naive Bayes Regressor}
+
+The Naive Bayes Regressor, and in particular, Bayesian Ridge, is the model used for regression in this case.
+
+% COMMENT copied FROM BayesianRidge source code
+% https://github.com/scikit-learn/scikit-learn/blob/9aaed4987/sklearn/linear_model/_bayes.py#L24
+%
+%     There exist several strategies to perform Bayesian ridge regression. This 
+%     implementation is based on the algorithm described in Appendix A of
+%     (Tipping, 2001) where updates of the regularization parameters are done as
+%     suggested in (MacKay, 1992). Note that according to A New
+%     View of Automatic Relevance Determination (Wipf and Nagarajan, 2008) these
+%     update rules do not guarantee that the marginal likelihood is increasing
+%     between two consecutive iterations of the optimization.
+
+The training parameters are listed in table \ref{table_bayesian_training_params}.
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{174pt}
+      \caption{Training parameters for XGBoost regressor}\label{table_bayesian_training_params}%
+      \begin{tabular}{@{}llll@{}}
+      \toprule
+      Parameter                     & Value \\
+      \midrule
+      Maximum iterations            & 300   \\
+      Tolerance\footnotemark[1]     & \num{1e-03} \\
+      alpha 1                       & \num{1e-06} \\
+      alpha 2                       & \num{1e-06} \\
+      lambda 1                      & \num{1e-06} \\
+      lambda 2                      & \num{1e-06} \\
+      \botrule
+      \end{tabular}
+      \footnotetext[1]{Tolerance for the stopping criteria.}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+\subsection{Fine-tuned Transformer}
+
+\textcolor{red}{TODO}
 

### 2023-04-06 paper/main.tex
analyzing results for paper
writing

@@
 
+The mean and standard deviation for each variable are as follows:
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Audio features description}\label{audio_features_stats}%
+      \begin{tabular}{@{}lll@{}}
+      \toprule
+      Feature           & $\mu$ & $\sigma$ \\
+      \midrule
+      Danceability      & 0.599  & 0.193  \\
+      Energy            & 0.631  & 0.233  \\
+      Acousticness      & 0.221  & 0.302  \\
+      Instrumentalness  & 0.514  & 0.382  \\
+      Valence           & 0.435  & 0.279  \\
+      \botrule
+      \end{tabular}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+Figure \ref{fig:audio_features_distribution} represents the density function of each variable.
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=\textwidth]{images/features-distribution.png}
+      \caption{Distribution of audio features}
+      \label{fig:audio_features_distribution}
+\end{figure}
+
+We can compare our single-user data with a bigger audio features dataset https://www.kaggle.com/datasets/tomigelo/spotify-audio-features.
+This dataset shows similar data distributions.
+
+I have posted a question on the Spotify Developer forum, relative to the release of a dataset 
+https://community.spotify.com/t5/Spotify-for-Developers/May-I-publish-an-open-source-dataset-that-contains-Spotify-Audio/td-p/5529259
+
 \subsection{Formatting Input Data for Predicting From Last.fm Tags}
@@
 
+\subsection{Experiments Execution and Results}
+
+The experiments:
+
+\begin{table}[h]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Experiment results. Cells values correspond to the RMSE value.}\label{experiment_results}%
+      \begin{tabular}{@{}llllll@{}}
+      \toprule
+      Experiment                           & Danceability    & Acousticness    & Energy    & Valence    & Instrumentalness \\
+      \midrule
+      $Base_token_{weight}$
+      $XGBoost_Tags_{energy}$                        & 300   \\
+      XGBoost - $y_{energy}$           & 300   \\
+      XGBoost - Tokens - Weight Repeat     & 300   \\
+      \botrule
+      \end{tabular}
+      \footnotetext[1]{Tolerance for the stopping criteria.}
+      \end{minipage}
+      \end{center}
+\end{table}
+
+\subsubsection{Results for Tabular Data Models}
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=0.5\textwidth]{images/rmse_by_model_and_limit.png}
+      \caption{RMSE mean and standard deviation by model and tags/tokens limit.}
+      \label{fig:rmse_by_model_and_limit}
+\end{figure}
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=0.7\textwidth]{images/rmse_by_model_and_feature.png}
+      \caption{RMSE mean and standard deviation by model and audio feature.}
+      \label{fig:rmse_by_model_and_feature}
+\end{figure}
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=0.7\textwidth]{images/rmse_by_model_and_input.png}
+      \caption{RMSE mean and standard deviation by model and input type (tag probablities or tokens).}
+      \label{fig:rmse_by_model_and_input}
+\end{figure}
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=0.7\textwidth]{images/rmse_by_feature_and_input.png}
+      \caption{RMSE mean and standard deviation by audio feature and input type (tag probablities or tokens).}
+      \label{fig:rmse_by_feature_and_input}
+\end{figure}
+
 \section{Conclusions}

### 2023-04-06 paper/main.tex
adding state of the art

@@
 
-Music information retrieval (MIR) is a field that focuses on the analysis,
+Music informationn retrieval (MIR) is a field that focuses on the analysis,
 processing and knowledge discovery of latent information in music pieces \cite{ramirez2020machine}.
@@
 
-% https://core.ac.uk/download/pdf/234915638.pdf
+In the last decade, Last.fm tags have been a popular source of metadata for MIR tasks.
+Last.fm tags can contain information the genre, mood, and style of music,
+and might be use to characterize certain features of a music piece.
 
-% https://sciendo.com/pdf/10.2478/ausi-2018-0009
+Last.fm tags can be useful when the audio signal available,
+for example, due to copyright limitations.
 
-% https://archives.ismir.net/ismir2009/paper/000018.pdf
+A number of studies have explored the use of Last.fm tags in MIR,
+and have shown promising results in predicting various audio features.
+% For example, tags such as "acoustic", "live", and "instrumental" have been
+% shown to be good predictors of the energy and danceability of a track.
+% Similarly, tags such as "happy," "sad," and "angry" have been used
+% to predict the valence and arousal of a track.
 
-Additionally, the Spotify audio features \dots
+Researchers have used Last.fm tags.
+For example, Laurier et al. analyzed how Last.fm tags categorize mood.
+In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.
 
-In general, these studies confirm the posibility of extracting knowledge from Last.fm tags.
+For example, {\c{C}}ano and Morisio discuss the process they follow to create a dataset of music
+lyrics annotated with Last.fm.
+In the creation process, they conclude that Last.fm tags are mostly related to music genre
+and positive moods \cite{ccano2017music}.
+In a similar direction, Bod{\'o} and Szil{\'a}gyi generated a dataset for lyrics genre classification
+by combining the Last.fm with MusicBrainz data \cite{bodo2018connecting}.
+MusicBrainz \footnotemark[1] is an online database of music editorial metadata.
+\footnote[1]{https://musicbrainz.org/}
+
+The Last.fm data has been the most widely used Last.fm dataset\footnotemark[2] in research.
+This dataset is a complementary dataset of the Million Song Dataset (MSD) \cite{Bertin-Mahieux2011}.
+
+\footnotetext[2]{Last.fm dataset, the official song tags and song similarity collection for the Million Song
+Dataset, available at: http://millionsongdataset.com/lastfm.}
+
+Additionally, the Spotify audio features have been used in multiple studies.
+Historically, these features were also called "EchoNest audio features".
+Echonest was an online platform was was later acquired by Spotify.
+% https://en.wikipedia.org/wiki/The_Echo_Nest
+
+Wang and Horv{\'a}t use audio features to study differences between male and female artists \cite{wang2019gender}
+
+In general, these studies confirm the possibility of extracting knowledge from Last.fm tags.
 To the best of our knowledge, no studies have addressed the problem of audio features regression, based solely on Last.fm tags.
+
+Jamdar et al. used EchoNest audio features, combined with lyrics data to classify songs into emotion tags.
+These classes were first defined based on a Last.fm tags emotion mapping \cite{jamdar2015emotion}.
+
+Similarly, Non-negative Matrix Factorization was applied in combination with EchoNest audio features
+for song recommendations \cite{benzi2016song}.
+
+P4kxspotify is a publicly available dataset that combines music review texts with Spotify audio features.
+The dataset creators argue that, although the terms of service prohibits scraping, their work is ethical \cite{pinter2020p4kxspotify}.
+
+In general, Spotify audio features have been used as predictive variables.
+We, to the best of our knowledge, are unaware of students that uses these features as target variables.
+
+\subsubsection{Machine Learning Models}
+
+Several studies have explored the use of machine learning
+models to predict audio features from audio metadata, such as Last.fm tags.
+% which studies use Last.fm tags as parameters?
+
+The experiments conduced in this study used two classical regression models,
+Boosted tree regressor and Bayesian ridge regressor.
+These two machine learning models that have been widely used in MIR. % is this true? (refs)
+
+Boosted tree regressor is a decision tree-based model that sequentially
+adds weak learners to the model to improve its performance.
+Bayesian ridge regressor, on the other hand,
+is a probabilistic model that uses Bayesian inference to estimate
+the parameters of the model.
+
+Additionally, language models, such as GTP-2 have shown promising results in various
+natural language processing (NLP) and generation.
+
+In this particular study, GTP-2 has been fine-tuned for regression tasks,
+and has shown good performance in predicting audio features from metadata.
+
+This paper aims to apply the boosted tree regressor,
+ Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
+The experiments compare the performance of these models and evaluate their effectiveness
+in predicting various audio features.
+The results of these experiments will provide insights into
+the use of different machine learning models for MIR tasks
+and can have practical applications
+in music recommendation systems and genre or mood recognition.
 

### 2023-04-24 paper/main.tex
working on intro

@@
 
-Music informationn retrieval (MIR) is a field that focuses on the analysis,
-processing and knowledge discovery of latent information in music pieces \cite{ramirez2020machine}.
+Music information retrieval (MIR) is an interdisciplinary research field that encompasses the extraction,
+processing, and knowledge discovery of information contained in music.
+MIR research intersects with other fields, such as computer science, signal processing, musicology, and sociology.
+The field covers applications in recommendation systems, music classification,
+music source separation, and music generation, among others \cite{ramirez2020machine}.
 
-Many of the challenges in this field use audio features, such as valence and energy, to predict other music-related aspects.
+MIR applications often attempt to extract information from the music audio signal,
+but analyzing associated metadata is also a common practice.
+Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based representations.
+Associated metadata, such as editorial information, lyrics, or user-generated content is usually in text format.
+However, metadata can be available as images of videos, for example, when analyzing album artwork, or music videos.
+
+Depending on the specific MIR application, researchers or practitioners expect different output values.
+For example, an application that extracts audio features might return values for the tempo, the key, or the sample rate.
+In more complex applications, where, for example, machine learning is used, applications might return the estimated emotion that a track produces on a listener, or the predicted music genre of a particular track.
+
+Many of the challenges in this field use audio features to predict other music-related aspects.
+Examples of these features are 
 For example, some studies have been using these features to predict the probably of a track being a hit [REF].

### 2023-05-02 paper/main.tex
first "readable" version of the intro

@@
 but analyzing associated metadata is also a common practice.
-Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based representations.
+Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based signal representations or hand-crafted audio features.
 Associated metadata, such as editorial information, lyrics, or user-generated content is usually in text format.
@@
 For example, an application that extracts audio features might return values for the tempo, the key, or the sample rate.
-In more complex applications, where, for example, machine learning is used, applications might return the estimated emotion that a track produces on a listener, or the predicted music genre of a particular track.
+In more complex applications, where, for example, machine learning is used,
+applications might return the estimated emotion that a track produces on a listener, or the predicted music genre of a particular track.
 
-Many of the challenges in this field use audio features to predict other music-related aspects.
-Examples of these features are 
-For example, some studies have been using these features to predict the probably of a track being a hit [REF].
+Among potentially useful input and output values, research has proved Spotify audio features and Last.fm tags to be significant values to characterize music.
+Spotify audio features capture high-level information about the music signal.
+Examples of these values are energy, danceability, or valence, among others.
+Last.fm tags are text labels that users associate to particular tracks, artists or albums in the Last.fm website.
 
-To make these predictions, researchers have been using machine learning models, specially over the past few years.
-Previously, researchers commonly used hand-crafted audio computational models to perform these tasks.
+Both types of values have been used as input values mostly for classification tasks, such as music genre recognition,
+where, given Spotify features or Last.fm tasks, the model estimates the music genres of a particular track.
+However, these values have not been used as target values by previous research, to the best of our knowledge.
+In particular, the focus of this article is on predicting Spotify audio features, given a set of Last.fm tags.
 
-But what if we tried to use these audio features as the predicted values? We would need a set of predictor variables to
-acurrately predict the value of audio features.
-This premise is the core concept of this article.
+By predicting Spotify audio features, we explore the relationship between subjective perception and concrete musical features.
+This approach might help to identify patterns and hidden correlations between how music is percevied, consumed and discovered.
 
-The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
-Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
-This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
+Additionally, predicted audio features could be used in recommendation systems to provide users with explainable recommendations.
+Music recommendations,are not always easy to interpret from the perspective of the listener.
+Users often get recommendations without meaningful explanations or justifications.
+Therefore, by using predicted Spotify features as basis for recommendation, we can explain users why the algorithm suggests a particular track.
+This process could be part of an explainable recommendation pipeline, where users enter a set of tags,
+and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
+
+% The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
+% Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
+% This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
 
@@
 
+In general, we believe that this novel approach that has the potential to benefit both listeners and researchers.
+By combining subjective user-generated tags with objective audio features,
+we can gain new insights into the complex relationship between perception and audio signal in music.
+
+Our approach also presents limitations.
+One limitation is the assumption of a strong relationship between Last.fm tags and Spotify features, which may not be true in all cases.
+Future work could explore other sources of input values, possibly related to the user context, to improve the accuracy of the predictions.
+
 \textcolor{red}{TODO}

### 2023-05-07 paper/main.tex
add reference
writing dataset creation section

@@
 This process could be part of an explainable recommendation pipeline, where users enter a set of tags,
-and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
+and they receive the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
 
@@
 
-\section{State of the Art}
+\section{Related Research}
 
-In recent years, researchers have used Last.fm tags in classification and regression tasks.
+In recent years, researchers have studied the use of Last.fm tags in classification and regression tasks.
 Several studies have used Last.fm to predict music sentiment or mood.
@@
 In the last decade, Last.fm tags have been a popular source of metadata for MIR tasks.
-Last.fm tags can contain information the genre, mood, and style of music,
+Last.fm tags can contain subjective information the genre, mood, and style of music,
 and might be use to characterize certain features of a music piece.
@@
 
+Spotify, one of the leaders in the music streaming industry...
+
 Additionally, the Spotify audio features have been used in multiple studies.
-Historically, these features were also called "EchoNest audio features".
-Echonest was an online platform was was later acquired by Spotify.
+Historically, these features were also called \emph{Echo Nest audio features}.
+Spotify adquired Echo Nest in 2014 and made the audio features available via the Spotify  API
+\footnote[3]{https://en.wikipedia.org/wiki/The\_Echo\_Nest}.
+Echonest was an online platform that was later acquired by Spotify.
 % https://en.wikipedia.org/wiki/The_Echo_Nest
@@
 
-\subsubsection{Machine Learning Models}
+While Spotify provides a description of the high-level audio features, how they compute or estimate these values is not publicly available.
+Panda and Redinho explore the use of Spotify high-level features applied to Music Emotion Recognition (MER) \cite{panda2021does}.
+In particular, they identify that the energy, valence, and acousticness values, provided by the Spotify API,
+are highly relevant for emotion classification.
+They also achieve better performance on MER models by using their own top-100 features, and they determine that,
+although these three Spotify features are relevant in terms of characterizing emotion, more features are needed for MER.
 
-Several studies have explored the use of machine learning
-models to predict audio features from audio metadata, such as Last.fm tags.
-% which studies use Last.fm tags as parameters?
+% TODO: move explanation to the experiment section.
+% \subsubsection{Machine Learning Models}
 
-The experiments conduced in this study used two classical regression models,
-Boosted tree regressor and Bayesian ridge regressor.
-These two machine learning models that have been widely used in MIR. % is this true? (refs)
+% Several studies have explored the use of machine learning
+% models to predict audio features from audio metadata, such as Last.fm tags.
+% % which studies use Last.fm tags as parameters?
 
-Boosted tree regressor is a decision tree-based model that sequentially
-adds weak learners to the model to improve its performance.
-Bayesian ridge regressor, on the other hand,
-is a probabilistic model that uses Bayesian inference to estimate
-the parameters of the model.
+% The experiments conduced in this study used two classical regression models,
+% Boosted tree regressor and Bayesian ridge regressor.
+% These two machine learning models that have been widely used in MIR. % is this true? (refs)
 
-Additionally, language models, such as GTP-2 have shown promising results in various
-natural language processing (NLP) and generation.
+% Boosted tree regressor is a decision tree-based model that sequentially
+% adds weak learners to the model to improve its performance.
+% Bayesian ridge regressor, on the other hand,
+% is a probabilistic model that uses Bayesian inference to estimate
+% the parameters of the model.
 
-In this particular study, GTP-2 has been fine-tuned for regression tasks,
-and has shown good performance in predicting audio features from metadata.
+% Additionally, language models, such as GTP-2 have shown promising results in various
+% natural language processing (NLP) and generation.
 
-This paper aims to apply the boosted tree regressor,
- Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
-The experiments compare the performance of these models and evaluate their effectiveness
-in predicting various audio features.
-The results of these experiments will provide insights into
-the use of different machine learning models for MIR tasks
-and can have practical applications
-in music recommendation systems and genre or mood recognition.
+% In this particular study, GTP-2 has been fine-tuned for regression tasks,
+% and has shown good performance in predicting audio features from metadata.
 
-\textcolor{red}{To be Continued \dots }
+% This paper aims to apply the boosted tree regressor,
+%  Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
+% The experiments compare the performance of these models and evaluate their effectiveness
+% in predicting various audio features.
+% The results of these experiments will provide insights into
+% the use of different machine learning models for MIR tasks
+% and can have practical applications
+% in music recommendation systems and genre or mood recognition.
+%
+%\textcolor{red}{To be Continued \dots }
 
@@
 
-Before conducting experiments on predicting audio features from tags, we constructed a dataset, by gathering the data from the Last.fm and Spotify APIs.
+% JAIME
+% I have posted a question on the Spotify Developer forum, relative to the release of a dataset
+% https://community.spotify.com/t5/Spotify-for-Developers/May-I-publish-an-open-source-dataset-that-contains-Spotify-Audio/td-p/5529259
+
+Before conducting experiments to predict audio features from tags,
+we constructed a dataset, by gathering the data from the Last.fm and Spotify APIs.
 
@@
 
+This work is scoped within our single-user research area \cite{ramirez2022user}.
+In this area, we explore the development of music recommender systems that
+characterize the music preferences and listening context only for a single user.
+By training our system with single-user data, we also explore the following question: Is it possible to train
+recommender systems, and in particular, user-centric systems, by using a single-user dataset?
+
+The user data for this work has been extracted from the
+listening history of the corresponding author user in Last.fm
+\footnote[4]{https://www.last.fm/user/jimmydj2000}.
+
 Similar to other intelligent systems, recommender systems
 must be trained, by using user preference data, to produce
-adequate recommendations. For our recommendation framework,
-we have leveraged the knowledge discovery potential
+suitable recommendations. For this study,
+we leverage the knowledge discovery potential
 of large historical listening logs, gathered from Last.fm.
 
-To characterize the preferences and context of the user, we
-have chosen to start with a simple scenario, where just data
-from a single user is available. By training our system with
-data from a single user, we also want to begin a discussion,
-given the following question: Is it possible to train
-recommender systems, and in particular, user-centric systems, by
-using a single-user dataset?
+Given the objective of our experiments,
+we created a dataset of Last.fm tags and Spotify audio features, indexed by track, by following these steps:
 
-To the best of our knowledge, research on user-centric rec-
-ommender systems has concentrated its efforts on explain-
-able AI Wang et al. [2019], and also user-centered evaluation
-of these systems Knijnenburg et al. [2012]. We also argue
-that recommender systems that exploit the preferences of a
-single user, or a reduced number of users, might as well be
-considered as user-centric models.
+\subsection{Last.fm Tags}
 
-\subsection{Last.fm}
+Last.fm is an online music service for users to keep track of their music listening habits.
+Last.fm is also an online community where users tag artists, albums, and tracks, according their own taste and perception.
 
-Last.fm is a online music service for uses to keep track of their music listenting habits.
-Last.fm can also be considered as an community where users tag artists, albums, and tracks, according the the own perception of the user.
-
-For nearly two decades, users have been contributing to Last.fm by tagging tracks with arbitrary text labels.
-These tags do not necessarily have to be single-worded.
-Users often use short sentences to define a song, such as `I like this track`, or `on the beach`.
-
-Community-contributed tags from the Last.fm API.
-These tags are text labels that Last.fm users assign
-to artists, albums, or tracks. Users apply these tags to
+Users apply these tags to
 categorize music from their own perspective, which
 means that tags do not fit into any structured ontology
-or data model. Tags can refer to aspects such as genre,
-emotion, or user context.
+or data model.
+Tags can refer to any aspect that users consider as a valid descriptor, such as genre,
+emotion, or user listenting context.
 
-\subsubsection{Last.fm Tags}
+For nearly two decades, users have been contributing to Last.fm by tagging tracks, albums, and artists with text labels.
+Although many of these descriptions are single-worded (e.g \emph{rock}, \emph{dance}, or \emph{happy}),
+users can also use short sentences to define a song, such as \emph{I like this track}, or \emph{on the beach}.
+
+\subsubsection{Last.fm Downloaded Data}
+
 Last.fm uses the term \emph{scrobble} to refer to a single track playback,
-in a particular moment. We have queried the Last.fm
-API to download the user{'}s scrobble logs, reported from 2007 to 2022.
+at a particular moment.
+We queried the Last.fm API to download the user{'}s scrobble logs, reported from 2007 to 2022.
 For each scrobble, we have gathered the following information:
@@
 \begin{itemize}
-     \item Track playback timestamp.
-     \item Track MusicBrainz Identifier (MBID), if exists.
+\item Playback timestamp
 \item Track name
@@
 \item Track tags. If the track does not have any tags assigned,
-      then artist tags have been used.
+then artist tags have been used
 \end{itemize}
 
-For each tag assigned to a track, or an artist, Last.fm includes
-a count property to indicate the popularity of the given tag for the track.
+For each tag-track mapping, Last.fm includes
+a \emph{count} value, which indicates the popularity of the given tag for the track.
 Last.fm normalizes this value in the 0-100 range, so the most popular tag for a track can have a
 count value of 100.
+For example, if \emph{piano} is the most popular tag for a track,
+then the track might be probably
+associated to the following tuple \verb|(piano, 100)|.
 
-Users normally listens to their favorite tracks many times,
-so the amount of individual tracks listened is much smaller
+Users normally listen to their favorite tracks several times,
+so the amount of unique tracks listened is much smaller
 than the number of track plays. In this case, the amount of
-individual tracks listened is about 20,000.
+individual tracks listened is about 20,000, and the number of scrobblings is, approximately, 90,000.
 
-The format is as follows:
+\subsection{Spotify Audio Features}
 
-\begin{verbatim}
-{
-      "artist - name": {
-            "eletronica": 100,
-            "rock": 80,
-            "pop": 45,
-            "jazz": 0,
-            "nu-jazz": 0,
-            "country": 0,
-            "soul": 0,
-      }
-}
-\end{verbatim}
+After gathering the listening history and track tags from Last.fm, and identifying the unique
+tracks that represent the user music collection, we
+collected Spotify audio features for each one of those tracks.
 
-\subsection{Spotify}
-
-After gathering Last.fm data and identifying the unique
-tracks that represent the user music collection, we have
-collected Spotify audio features. For each of these individ-
-ual tracks, we have downloaded the Spotify audio features
-specific to the given track.
-
-The Spotify audio features are numerical values that repre-
-sent high-level audio information computed from a specific
+The Spotify audio features are numerical values that represent high-level audio information computed from a specific
 track. These values characterize a track, musically speaking,
 by measuring relevant musical aspects.
-
-The features provided by the Spotify API are acousticness,
-danceability, duration\textunderscore ms, energy, instrumentalness, key,
-liveness, loudness, mode, speechiness, tempo, and valence.
-Table 1 describes these features. The reader can find further
-details about each feature in the Spotify API documentation
-4.
-
-A small portion of the tracks do not have features available in
-Spotify, so they have been filtered out from our experiments.
-
-After filtering songs that miss Last.fm tags or Spotify audio features,
-our dataset contains 14009 samples.
-
-\emph{Track audio features from the Spotify API}. These are
-attributes computed from the audio themselves. They
-are a way to describe music by using numerical values.
-For example, a danceability attribute of 0.95 means
+For example, a danceability value of 0.95 means
 that a particular song is highly suitable for dancing.
 
-\subsection{Last.fm Tags Representations for Training}
+The features provided by the Spotify API are listed in
+Table \ref{table:spotify-features}.
+The reader can find further details about each feature in the Spotify API documentation
+\footnote[5]{
+      \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
+}.
 
-\subsubsection{Tabular}
-Each tag is a column and each cell contains the popularity value of a tag for a track.
-A cell is 0 if a tag is missing for a track.
+\begin{table}[h!]
+      \centering
+      \caption{Spotify audio features. These features provide high-level musical information about a track.} \label{table:spotify-features}
+      \begin{tabular}{p{0.3\linewidth}p{0.6\linewidth}}
+          \toprule
+          \bfseries \textbf{Feature name} & \textbf{Description} \\
+          \midrule
+          \textbf{acousticness} & The track is acoustic. From 0 to 1 \\
+          \textbf{danceability} & The track encourages (or is adequate for) dancing. From 0 to 1 \\
+          \textbf{duration\_ms}  &  Duration in milliseconds \\
+          \textbf{energy}  &  The track is perceived as energetic. From 0 to 1\\
+          \textbf{instrumentalness}  &  The track is instrumental. From 0 to 1 \\
+          \textbf{key}  &  Key categories encoded as integers. From C (0) to 11 \\
+          \textbf{liveness}  &  The audience is audible. From 0 to 1\\
+          \textbf{loudness}  &  In decibels. From -60 to 0 \\
+          \textbf{mode}  & Major (1) or minor (0) \\
+          \textbf{speechiness}  & Does the track contain speeches? From 0 to 1 \\
+          \textbf{tempo}  & In beats per minute (BPM) \\
+          \textbf{valence} & How happy is the track (BPM). From 0 to 1 \\
+          \bottomrule
+      \end{tabular}
+  \end{table}
 
-The number of columns is limited to the top-K tags.
+The Spotify API failed to provide audio features for a portion of the tracks.
+These tracks were filtered out from our experiments.
+After filtering tracks that miss Last.fm tags or Spotify audio features,
+our dataset contains 14,009 samples.
 
-\subsubsection{Tabular Tokens}
+Considering that the data was gathered from a single user,
+we explored the data to verify that the distribution of the Spotify audio features
+was comparable to other Spotify datasets.
+In particular, we verified that the distribution of the features,
+described in Table \ref{table:audio-features-stats} and Figure \ref{fig:audio-features-distribution},
+was comparable to the distribution of the \emph{Spotify Audio Features} Kaggle dataset
+\footnote[6]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
 
-Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
-To tokenize tags, we have used the GTP2 tokenizer.
-Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
-To \emph{stringify} the tags, we have concatenated tags with multiple strategies:
-
-\begin{itemize}
-      \item By including tag popularity: `rock 2, pop 1`.
-      \item By repeating tags based on popularity: `rock rock, pop`.
-      \item By ordering by popularity: `rock, pop`.
-\end{itemize}
-
-\subsection{Training Data By Track}
-
-When generating training data by track, the tabular formats present sparsity problems.
-
-For tabular representations, we need to defined a fixed set of columns as tags.
-For most of tracks, most columns are `0`.
-
-The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
-(e.g., m * n for an m * n matrix) is called the sparsity of the matrix
-
-The mean and standard deviation for each variable are as follows:
-
-\begin{table}[h]
+\begin{table}[h!]
       \begin{center}
       \begin{minipage}{\textwidth}
-      \caption{Audio features description}\label{audio_features_stats}%
+      \caption{Audio features description}\label{table:audio-features-stats}%
       \begin{tabular}{@{}lll@{}}
@@
 
-Figure \ref{fig:audio_features_distribution} represents the density function of each variable.
-
 \begin{figure}[h!]
@@
       \includegraphics[width=\textwidth]{images/features-distribution.png}
-      \caption{Distribution of audio features}
-      \label{fig:audio_features_distribution}
+      \caption{Distribution of audio features. These densities are similar to the values shown by the \emph{Spotify Audio Features} Kaggle dataset.}
+      \label{fig:audio-features-distribution}
 \end{figure}
 
-We can compare our single-user data with a bigger audio features dataset https://www.kaggle.com/datasets/tomigelo/spotify-audio-features.
-This dataset shows similar data distributions.
+\subsection{Filtering Missing Values}
 
-I have posted a question on the Spotify Developer forum, relative to the release of a dataset
-https://community.spotify.com/t5/Spotify-for-Developers/May-I-publish-an-open-source-dataset-that-contains-Spotify-Audio/td-p/5529259
+Not every track of the approximately 20,000 included in the Last.fm user{'} listening history
+included Last.fm tags and Spotify audio features.
+Tracks without Last.fm tags or without Spotify audio features were filtered out.
+As a result, of the originally 20,000 tracks included in the listening history,
+our dataset final size resulted in 14,009 samples.
 
-\subsection{Formatting Input Data for Predicting From Last.fm Tags}
+\subsection{Last.fm Tags Input Format}
 
-The input data passed to the XGBoost regressor is formatted in tabular format, as follows:
+Each individual sample in the dataset corresponds to a unique track,
+and contains the list of Last.fm tag-count tuples (e.g. \verb|[(electronic, 100), (dance, 45), ...]|)
+and the values of Spotify audio features.
+Before experimenting with machine learning models, we prepared the data in a number of different formats,
+each one suitable for specific models.
+
+\subsubsection{Tabular}
+Each tag is a column and each cell contains the count value of a tag for a track.
+A cell is 0 if a tag is missing for a track.
+
+The number of columns is limited to the top-K tags.
+Counting the total amount of Last.fm tags in the user collection resulted, initially, in more than five million tags.
+We quickly confirmed that building a tabular data set, in which every row contains millions of columns (Last.fm tags)
+was doable, but presented scalability problems.
+Therefore, we reduced the number of tags by picking a subset of the most relevant tags.
+
+The reduction algorithm is simple: calculate a weighted sum of all the tags appearances and pick the top 1000.
+The sum is weighted be- cause we use the count attribute.
+This attribute is present in every Last.fm track-tag association and provides a measure of the strength of a particular tag in a specific track.
+Note that this reduction is an initial approach, which, similar to other phases, can be extended or improved in the future.
+For this particular case, dimensionality reduction algorithms, such as PCA, are good candidates for forthcoming iterations of this work.
+
+The input data passed is formatted in tabular format, as follows:
 
@@
 
-\subsection{Formatting Input Data for Predicting From Tokens}
+When generating training data by track, the tabular formats present sparsity problems.
+
+For tabular representations, we need to defined a fixed set of columns as tags.
+For most of tracks, most columns are 0.
+
+The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
+(e.g., m * n for an m * n matrix) is called the sparsity of the matrix.
+
+\subsubsection{Tabular Tokens}
+
+Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
+To tokenize tags, we have used the GTP2 tokenizer.
+Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
+To \emph{stringify} the tags, we have concatenated tags with multiple strategies:
+
+\begin{itemize}
+      \item By including tag popularity: `rock 2, pop 1`.
+      \item By repeating tags based on popularity: `rock rock, pop`.
+      \item By ordering by popularity: `rock, pop`.
+\end{itemize}
 
@@
 
-\subsection{Formatting Input Data for Predicting From Text}
+\subsection{Text}
 

### 2023-05-11 paper/main.tex
first vesrion of results table

@@
 
-Not every track of the approximately 20,000 included in the Last.fm user{'} listening history
+Not every track of the approximately unique tracks 20,000 of the Last.fm user{'} listening history
 included Last.fm tags and Spotify audio features.
@@
 our dataset final size resulted in 14,009 samples.
+
+\section{Experiments}
+
+We trained commonly used machine learning models to predict an audio feature, given the set of tags for a particular track.
+
+\begin{itemize}
+      \item Boosted tree regressor \cite{xgboost}
+      \item Naive Bayes Regressor \cite{bayesian}
+      \item Fine-tuned GPT-2 model
+\end{itemize}
+
+The preceding models require specific input formats.
+Therefore, we tested different input formats.
 
@@
 \subsubsection{Tabular}
-Each tag is a column and each cell contains the count value of a tag for a track.
+With this format, the Last.fm tags are represented as a table.
+Each tag is defined by a column and each cell contains the count value of a tag for a track.
 A cell is 0 if a tag is missing for a track.
@@
 
-\section{Experiments}
-
-We trained a commonly used machine learning models to predict an audio feature, given the set of tags for a particular track.
-
-\begin{itemize}
-      \item Boosted tree regressor \cite{xgboost}
-      \item Naive Bayes Regressor \cite{bayesian}
-      \item Fine-tuned GPT-2 model
-\end{itemize}
-
 \subsection{Boosted Tree Regressor}
@@
 
-\textcolor{red}{TODO}
+Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
+such as text generation and question answering.
+They were first introduced in 2017 by paper [REF: "Attention Is All You Need"]
+and have become one of the most popular models due to their ability
+to handle long-range dependencies and process variable-length inputs.
+
+One popular variant of transformers is the Generative Pre-trained Transformer 2 (GPT-2),
+which was introduced by OpenAI in 2019.
+GPT-2 is a language model that has been trained on a large corpus of text,
+and can generate coherent and fluent text that resembles human-written language.
+GPT-2 has achieved remarkable results in a wide range of natural language processing tasks, including text generation, machine translation, and question answering.
+
+In this study, we have used GPT-2 as a regressor to predict Spotify audio features from Last.fm tags.
+The basic idea behind this approach is to feed a string of concatenated Last.fm tags as input to the GPT-2 model, and then use the model's output as the predicted value of a specific Spotify audio feature.
+By training the GPT-2 model on a large dataset of Last.fm tags and corresponding Spotify audio features,
+we aim to learn the complex relationships between these two types of data and to use this knowledge to evaluate the prediction accuracy of Spotify audio features based on Last.fm tags.
+
+To train the GPT-2 model as a regressor, we used a supervised learning approach.
+Specifically, we collected a large dataset of Last.fm tags and corresponding Spotify audio features, and used this dataset to train the GPT-2 model to predict Spotify audio features from Last.fm tags.
+We used a mean squared error loss function to optimize the model's performance during training, and we also employed techniques such as early stopping and learning rate scheduling to prevent overfitting and improve generalization performance.
+\textcolor{red}{Under development...}
 
@@
 
-The experiments:
+Table \ref{table:experiment_results} summaries the experiment results.
+The table provides RMSE values for each experiment.
 
-\begin{table}[h]
+\begin{table}[h!]
       \begin{center}
       \begin{minipage}{\textwidth}
-      \caption{Experiment results. Cells values correspond to the RMSE value.}\label{experiment_results}%
-      \begin{tabular}{@{}llllll@{}}
+      \caption{Experiment results. Cells values correspond to the RMSE value.}\label{table:experiment_results}%
+      \begin{tabular}{@{}lllllll@{}}
       \toprule
-      Experiment                           & Danceability    & Acousticness    & Energy    & Valence    & Instrumentalness \\
+      M         & Input format                        & Danceab...       & Acoustic...    & Energy    & Valence    & Instrumen... \\
       \midrule
-      $Base_token_{weight}$
-      $XGBoost_Tags_{energy}$                        & 300   \\
-      XGBoost - $y_{energy}$           & 300   \\
-      XGBoost - Tokens - Weight Repeat     & 300   \\
+      Base      &                                     & 0.276            & 0.438           & 0.329     & 0.395      & 0.541          \\
+      Bayes     & 100 tags\footnotemark[1]            & 0.159            & 0.261           & 0.197     & 0.243      & 0.307          \\
+      Tree      & 100 tags                            & 0.154            & 0.257           & 0.188     & 0.240      & 0.302          \\
+      Bayes     & 1000 tags                           & 0.153            & 0.253           & 0.190     & 0.237      & 0.299          \\
+      Tree      & 1000 tags                           & 0.149            & 0.249           & 0.184     & 0.236      & 0.291          \\
+      Bayes     & 10000 tags                          & 0.152            & 0.251           & 0.189     & 0.236      & 0.297          \\
+      Tree      & 10000 tags                          & 0.148            & 0.250           & 0.181     & 0.236      & 0.290          \\
+      Bayes     & 100 tokens D\footnotemark[2]        & 0.307            & 0.307           & 0.238     & 0.281      & 0.383          \\
+      Tree      & 100 tokens D                        & 0.274            & 0.274           & 0.212     & 0.256      & 0.330          \\
+      Bayes     & 1000 tokens D                       & 0.201            & 0.315           & 0.249     & 0.248      & 0.399          \\
+      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215     & 0.220      & 0.339          \\
+      Bayes     & 10000 tokens D                      & 0.359            & 0.507           & 0.394     & 0.479      & 0.613          \\
+      Tree      & 10000 tokens D                      & 0.172            & 0.276           & 0.217     & 0.266      & 0.342          \\
+      Bayes     & 100 tokens O\footnotemark[3]        & 0.191            & 0.305           & 0.237     & 0.282      & 0.376          \\
+      Tree      & 100 tokens O                        & 0.179            & 0.294           & 0.225     & 0.271      & 0.350          \\
+      Bayes     & 1000 tokens O                       & 0.237            & 0.343           & 0.276     & 0.339      & 0.428          \\
+      Tree      & 1000 tokens O                       & 0.182            & 0.294           & 0.224     & 0.270      & 0.353          \\
+      Bayes     & 10000 tokens O                      & 0.237            & 0.343           & 0.276     & 0.339      & 0.428          \\
+      Tree      & 10000 tokens O                      & 0.182            & 0.294           & 0.225     & 0.267      & 0.353          \\
+      Bayes     & 100 tokens TC\footnotemark[4]       & 0.191            & 0.304           & 0.236     & 0.281      & 0.380          \\
+      Tree      & 100 tokens TC                       & 0.172            & 0.280           & 0.211     & 0.262      & 0.342          \\
+      Bayes     & 1000 tokens TC                      & 0.202            & 0.320           & 0.247     & 0.248      & 0.404          \\
+      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215     & 0.220      & 0.344          \\
+      Bayes     & 10000 tokens TC                     & 0.234            & 0.341           & 0.274     & 0.321      & 0.430          \\
+      Tree      & 10000 tokens TC                     & 0.175            & 0.281           & 0.214     & 0.270      & 0.345          \\
+      .....     & .............                       & .....            & .....           & .....     & .....      & .....          \\
+      GPT       & Duplicated\footnotemark[2]          & 0.157            & ...             & ...       & ...        & ...            \\
+      GPT       & Ordered\footnotemark[3]             & 0.149            & ...             & ...       & ...        & ...            \\
+      GPT       & Tags,Counts\footnotemark[4]         & \textbf{0.145}   & ...             & ...       & ...        & ...            \\
       \botrule
       \end{tabular}
-      \footnotetext[1]{Tolerance for the stopping criteria.}
+      \footnotetext[1]{Tags in tabular format. Given $(rock, 3)$, the cell in the $rock$ column contains \verb*|3|.}
+      \footnotetext[2]{Duplicated tokens in tabular format. Tags $(rock, 3), (pop, 2)$, are converted to the \verb*|"rock, rock, rock, pop, pop"| string, 
+      which a tokenizer converts to the list of input tokens (e.g. \verb*|[101,1005,16588,1005,2531, ...]|).
+      These tokens are passed to the model in tabular format. Columns are $token_1$, $token_2$, $...$, $token_N$.}
+      \footnotetext[3]{Ordered tokens in tabular format. Tags $(rock, 3), (pop, 2)$, are converted to the \verb*|"rock, pop"| string,
+      which a tokenizer converts to the list of input tokens (e.g. \verb*|[101,1005,16588,1005,2531, ...]|).
+      These tokens are passed to the model in tabular format. Columns are $token_1$, $token_2$, $...$, $token_N$.}
+      \footnotetext[4]{Tokens in tabular format from tags and counts. Tags $(rock, 3), (pop, 2)$, are converted to the \verb*|"'rock' 3, 'pop' 2"| string,
+      which a tokenizer converts to the list of input tokens (e.g. \verb*|[101,1005,16588,1005,2531, ...]|).
+      These tokens are passed to the model in tabular format. Columns are $token_1$, $token_2$, $...$, $token_N$.}
+      \footnotetext[2]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, rock, rock, pop, pop"|.}
+      \footnotetext[3]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, pop"|.}
+      \footnotetext[4]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"'rock' 3, 'pop' 2"|.}
       \end{minipage}

### 2023-05-11 paper/main.tex
improved results table and more details about kaggle dataset

@@
 described in Table \ref{table:audio-features-stats} and Figure \ref{fig:audio-features-distribution},
-was comparable to the distribution of the \emph{Spotify Audio Features} Kaggle dataset
+was comparable to the distribution of the \emph{Spotify Audio Features} Kaggle dataset.
+This dataset contains more tha 116,000 unique tracks, and includes audio features for each track.
+Our dataset present similar $\mu$ and $\sigma$ values as the Spotify Audio Features Kaggle dataset,
+as illustrated in Table \ref{table:audio-features-stats-kaggle} and Figure \ref{fig:audio-features-distribution-kaggle}.
 \footnote[6]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
+
+% Another dataset created from the Spotify API and published to Kaggle:
+% https://www.kaggle.com/datasets/andrewmvd/spotify-playlists
 
@@
       \midrule
-      Danceability      & 0.599  & 0.193  \\
-      Energy            & 0.631  & 0.233  \\
-      Acousticness      & 0.221  & 0.302  \\
-      Instrumentalness  & 0.514  & 0.382  \\
-      Valence           & 0.435  & 0.279  \\
+      Danceability      & 0.60  & 0.19  \\
+      Energy            & 0.63  & 0.23  \\
+      Acousticness      & 0.22  & 0.30  \\
+      Instrumentalness  & 0.51  & 0.38  \\
+      Valence           & 0.44  & 0.28  \\
+      \botrule
+      \end{tabular}
+      \end{minipage}
+      \end{center}
+\end{table}
+\begin{table}[h!]
+      \begin{center}
+      \begin{minipage}{\textwidth}
+      \caption{Audio features description of \emph{Spotify Audio Features} Kaggle dataset}\label{table:audio-features-stats-kaggle}%
+      \begin{tabular}{@{}lll@{}}
+      \toprule
+      Feature           & $\mu$ & $\sigma$ \\
+      \midrule
+      Danceability      & 0.58   & 0.19  \\
+      Energy            & 0.57  & 0.26  \\
+      Acousticness      & 0.34  & 0.25  \\
+      Instrumentalness  & 0.22  & 0.36  \\
+      Valence           & 0.44  & 0.26  \\
       \botrule
@@
       \includegraphics[width=\textwidth]{images/features-distribution.png}
-      \caption{Distribution of audio features. These densities are similar to the values shown by the \emph{Spotify Audio Features} Kaggle dataset.}
+      \caption{Distribution of audio features in our single-user dataset.}
       \label{fig:audio-features-distribution}
+\end{figure}
+
+\begin{figure}[h!]
+      \centering
+      \includegraphics[width=\textwidth]{images/feature-distribution-kaggle.png}
+      \caption{Distribution of audio features in the \emph{Spotify Audio Features} Kaggle dataset.}
+      \label{fig:audio-features-distribution-kaggle}
 \end{figure}
@@
       Base      &                                     & 0.276            & 0.438           & 0.329          & 0.395           & 0.541          \\
+      \midrule
       Bayes     & 100 tags\footnotemark[1]            & 0.159            & 0.261           & 0.197          & 0.243           & 0.307          \\
+      Bayes     & 1000 tags                           & 0.153            & 0.253           & 0.190          & 0.237           & 0.299          \\
+      Bayes     & 10000 tags                          &\textbf{0.152}    &\textbf{0.251}   &\textbf{0.189}  &\textbf{0.236}   &\textbf{0.297}  \\
+      Bayes     & 100 tokens D\footnotemark[2]        & 0.307            & 0.307           & 0.238          & 0.281           & 0.383          \\
+      Bayes     & 10000 tokens D                      & 0.359            & 0.507           & 0.394          & 0.479           & 0.613          \\
+      Bayes     & 1000 tokens D                       & 0.201            & 0.315           & 0.249          & 0.248           & 0.399          \\
+      Bayes     & 100 tokens O\footnotemark[3]        & 0.191            & 0.305           & 0.237          & 0.282           & 0.376          \\
+      Bayes     & 1000 tokens O                       & 0.237            & 0.343           & 0.276          & 0.339           & 0.428          \\
+      Bayes     & 10000 tokens O                      & 0.237            & 0.343           & 0.276          & 0.339           & 0.428          \\
+      Bayes     & 100 tokens TC\footnotemark[4]       & 0.191            & 0.304           & 0.236          & 0.281           & 0.380          \\
+      Bayes     & 1000 tokens TC                      & 0.202            & 0.320           & 0.247          & 0.248           & 0.404          \\
+      Bayes     & 10000 tokens TC                     & 0.234            & 0.341           & 0.274          & 0.321           & 0.430          \\
+      \midrule
       Tree      & 100 tags                            & 0.154            & 0.257           & 0.188          & 0.240           & 0.302          \\
-      Bayes     & 1000 tags                           & 0.153            & 0.253           & 0.190     & 0.237      & 0.299          \\
-      Tree      & 1000 tags                           & 0.149            & 0.249           & 0.184     & 0.236      & 0.291          \\
-      Bayes     & 10000 tags                          & 0.152            & 0.251           & 0.189     & 0.236      & 0.297          \\
-      Tree      & 10000 tags                          & 0.148            & 0.250           & 0.181     & 0.236      & 0.290          \\
-      Bayes     & 100 tokens D\footnotemark[2]        & 0.307            & 0.307           & 0.238     & 0.281      & 0.383          \\
+      Tree      & 1000 tags                           & 0.149            &\textbf{0.249}   & 0.184          & 0.236           & 0.291          \\
+      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  &\textbf{0.235}   &\textbf{0.290}  \\
       Tree      & 100 tokens D                        & 0.274            & 0.274           & 0.212          & 0.256           & 0.330          \\
-      Bayes     & 1000 tokens D                       & 0.201            & 0.315           & 0.249     & 0.248      & 0.399          \\
       Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & 0.220           & 0.339          \\
-      Bayes     & 10000 tokens D                      & 0.359            & 0.507           & 0.394     & 0.479      & 0.613          \\
       Tree      & 10000 tokens D                      & 0.172            & 0.276           & 0.217          & 0.266           & 0.342          \\
-      Bayes     & 100 tokens O\footnotemark[3]        & 0.191            & 0.305           & 0.237     & 0.282      & 0.376          \\
       Tree      & 100 tokens O                        & 0.179            & 0.294           & 0.225          & 0.271           & 0.350          \\
-      Bayes     & 1000 tokens O                       & 0.237            & 0.343           & 0.276     & 0.339      & 0.428          \\
       Tree      & 1000 tokens O                       & 0.182            & 0.294           & 0.224          & 0.270           & 0.353          \\
-      Bayes     & 10000 tokens O                      & 0.237            & 0.343           & 0.276     & 0.339      & 0.428          \\
       Tree      & 10000 tokens O                      & 0.182            & 0.294           & 0.225          & 0.267           & 0.353          \\
-      Bayes     & 100 tokens TC\footnotemark[4]       & 0.191            & 0.304           & 0.236     & 0.281      & 0.380          \\
       Tree      & 100 tokens TC                       & 0.172            & 0.280           & 0.211          & 0.262           & 0.342          \\
-      Bayes     & 1000 tokens TC                      & 0.202            & 0.320           & 0.247     & 0.248      & 0.404          \\
       Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & 0.220           & 0.344          \\
-      Bayes     & 10000 tokens TC                     & 0.234            & 0.341           & 0.274     & 0.321      & 0.430          \\
       Tree      & 10000 tokens TC                     & 0.175            & 0.281           & 0.214          & 0.270           & 0.345          \\
-      .....     & .............                       & .....            & .....           & .....     & .....      & .....          \\
-      GPT       & Duplicated\footnotemark[2]          & 0.157            & ...             & ...       & ...        & ...            \\
-      GPT       & Ordered\footnotemark[3]             & 0.149            & ...             & ...       & ...        & ...            \\
-      GPT       & Tags,Counts\footnotemark[4]         & \textbf{0.145}   & ...             & ...       & ...        & ...            \\
+      \midrule
+      GPT       & Duplicated\footnotemark[5]          & 0.157            & ...             & ...            & ...             & ...            \\
+      GPT       & Ordered\footnotemark[6]             & 0.149            & ...             & ...            & ...             & ...            \\
+      GPT       & Tags,Counts\footnotemark[7]         & \textbf{0.145}   & ...             & ...            & ...             & ...            \\
+
       \botrule
@@
       These tokens are passed to the model in tabular format. Columns are $token_1$, $token_2$, $...$, $token_N$.}
-      \footnotetext[2]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, rock, rock, pop, pop"|.}
-      \footnotetext[3]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, pop"|.}
-      \footnotetext[4]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"'rock' 3, 'pop' 2"|.}
+      \footnotetext[5]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, rock, rock, pop, pop"|.}
+      \footnotetext[6]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"rock, pop"|.}
+      \footnotetext[7]{String. Given tags $(rock, 3), (pop, 2)$, input is formatted as \verb*|"'rock' 3, 'pop' 2"|.}
       \end{minipage}

### 2023-05-12 paper/main.tex
add comments
add missing results in table
improving input formatting section

@@
 
-\title[Predicting Audio Features with Last.fm Tags]{Predicting Audio Features with Last.fm Tags}
+\title[Predicting Spotify Audio Features from Last.fm Tags]{Predicting Spotify Audio Features from Last.fm Tags}
 
@@
 
-Both types of values have been used as input values mostly for classification tasks, such as music genre recognition,
+Both types of values have been used as input data mostly for classification tasks, such as music genre recognition,
 where, given Spotify features or Last.fm tasks, the model estimates the music genres of a particular track.
-However, these values have not been used as target values by previous research, to the best of our knowledge.
+Previous studies, however, have not used these values as target outputs, to the best of our knowledge.
 In particular, the focus of this article is on predicting Spotify audio features, given a set of Last.fm tags.
@@
 Additionally, predicted audio features could be used in recommendation systems to provide users with explainable recommendations.
-Music recommendations,are not always easy to interpret from the perspective of the listener.
+Music recommendations are tipically dificult to interpret from the perspective of the listener.
 Users often get recommendations without meaningful explanations or justifications.
@@
 
-In recent years, researchers have studied the use of Last.fm tags in classification and regression tasks.
-Several studies have used Last.fm to predict music sentiment or mood.
+\subsubsection{Last.fm Tags}
 
-In the last decade, Last.fm tags have been a popular source of metadata for MIR tasks.
-Last.fm tags can contain subjective information the genre, mood, and style of music,
-and might be use to characterize certain features of a music piece.
+In the last decade, researchers have studied the use of Last.fm tags in classification and regression tasks.
+Last.fm tags have been a popular source of metadata for MIR tasks,
+because they potentionally contain subjective information related to the genre, mood, and style of music,
+and might be used to characterize certain features of a music piece.
+Several studies have used Last.fm to predict music sentiment or mood, and to predict audio features.
 
-Last.fm tags can be useful when the audio signal available,
+Additionally, Last.fm tags constitute a useful source of input knownledge when the audio signal is not available,
 for example, due to copyright limitations.
 
-A number of studies have explored the use of Last.fm tags in MIR,
-and have shown promising results in predicting various audio features.
+% == Excerpt taken from one of our references vvvvvvv
 % For example, tags such as "acoustic", "live", and "instrumental" have been
@@
 
-Researchers have used Last.fm tags.
-For example, Laurier et al. analyzed how Last.fm tags categorize mood.
+For example, Laurier et al. analyze how Last.fm tags categorize mood.
 In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.
 
-For example, {\c{C}}ano and Morisio discuss the process they follow to create a dataset of music
-lyrics annotated with Last.fm.
+{\c{C}}ano and Morisio discuss how they create a dataset of music
+lyrics annotated with Last.fm tags.
 In the creation process, they conclude that Last.fm tags are mostly related to music genre
 and positive moods \cite{ccano2017music}.
-In a similar direction, Bod{\'o} and Szil{\'a}gyi generated a dataset for lyrics genre classification
+
+In a similar direction, Bod{\'o} and Szil{\'a}gyi generate a dataset for lyrics genre classification
 by combining the Last.fm with MusicBrainz data \cite{bodo2018connecting}.
-MusicBrainz \footnotemark[1] is an online database of music editorial metadata.
-\footnote[1]{https://musicbrainz.org/}
+MusicBrainz is an online database of music editorial metadata\footnote[1]{https://musicbrainz.org/}.
 
-The Last.fm data has been the most widely used Last.fm dataset\footnotemark[2] in research.
+So far, The \emph{Last.fm dataset} has been the most widely used Last.fm dataset\footnotemark[2] in research.
 This dataset is a complementary dataset of the Million Song Dataset (MSD) \cite{Bertin-Mahieux2011}.
@@
 
-Spotify, one of the leaders in the music streaming industry...
+In general, these studies confirm the possibility of extracting knowledge from Last.fm tags.
 
-Additionally, the Spotify audio features have been used in multiple studies.
-Historically, these features were also called \emph{Echo Nest audio features}.
-Spotify adquired Echo Nest in 2014 and made the audio features available via the Spotify  API
-\footnote[3]{https://en.wikipedia.org/wiki/The\_Echo\_Nest}.
-Echonest was an online platform that was later acquired by Spotify.
-% https://en.wikipedia.org/wiki/The_Echo_Nest
+\subsubsection{Spotify Audio Features}
 
-Wang and Horv{\'a}t use audio features to study differences between male and female artists \cite{wang2019gender}
+% Spotify, one of the leaders in the music streaming industry...
 
-In general, these studies confirm the possibility of extracting knowledge from Last.fm tags.
-To the best of our knowledge, no studies have addressed the problem of audio features regression, based solely on Last.fm tags.
+The Spotify audio features have been used in multiple studies.
+Historically, these features were called the \emph{Echo Nest audio features}.
+\emph{The Echo Nest} was an online music intelligence platform that provided users and clients with music analysis services.
+Among these services, the Echo Nest offered a database and an API to retrieve audio features for each of the tracks in the database \footnote[3]{https://en.wikipedia.org/wiki/The\_Echo\_Nest}.
+Spotify acquired the Echo Nest in 2014.
+As a result, the Echo Nest API was shut down and Spotify made these audio features available via the Spotify API.
+
+Nowadays, published research is available under the two terms.
+Whereas the most recent studies refer to these audio features as the Spotify audio features,
+earlier studies use the Echo Nest denomination.
+
+Regardless of the term used, the features remain the same.
+These features are a set of high-level descriptors, such as energy and danceability,
+which are related to the audio but also to the listeners perception.
+
+While Spotify provides a description of the audio features, how they compute or estimate these values is not publicly available.
+
+Wang and Horv{\'a}t use audio features to study differences between male and female artists \cite{wang2019gender}.
 
@@
 
-P4kxspotify is a publicly available dataset that combines music review texts with Spotify audio features.
-The dataset creators argue that, although the terms of service prohibits scraping, their work is ethical \cite{pinter2020p4kxspotify}.
-
-In general, Spotify audio features have been used as predictive variables.
-We, to the best of our knowledge, are unaware of students that uses these features as target variables.
-
-While Spotify provides a description of the high-level audio features, how they compute or estimate these values is not publicly available.
 Panda and Redinho explore the use of Spotify high-level features applied to Music Emotion Recognition (MER) \cite{panda2021does}.
@@
 although these three Spotify features are relevant in terms of characterizing emotion, more features are needed for MER.
+
+Publicly available datasets of Spotify audio features can be found online,
+as a result of open-source and research communities collecting data from the Spotify API and publishing the results.
+It is unclear, however, whether these published datasets violate the Spotify API terms of service.
+
+For example, \emph{P4kxspotify} is a publicly available dataset that combines music review texts with Spotify audio features.
+The dataset creators argue that, although the terms of service prohibits scraping, their work is ethical \cite{pinter2020p4kxspotify}.
+
+Another example is the \emph{Spotify Audio Features} Kaggle dataset\footnote[4]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
+This dataset contains more than 116,000 unique tracks, and includes audio features for each track.
+
+% Another dataset created from the Spotify API and published to Kaggle:
+% https://www.kaggle.com/datasets/andrewmvd/spotify-playlists
+
+In general, Spotify audio features have been used as predictive input variables.
+We, to the best of our knowledge, are unaware of studies that use these features as target variables,
+or studies that have addressed the problem of audio features regression, based solely on Last.fm tags.
 
@@
 
-This work is scoped within our single-user research area \cite{ramirez2022user}.
+This work is scoped within our single-user research line \cite{ramirez2022user}.
 In this area, we explore the development of music recommender systems that
 characterize the music preferences and listening context only for a single user.
-By training our system with single-user data, we also explore the following question: Is it possible to train
+By training our system with single-user data, we also raise the following question: Is it possible to train
 recommender systems, and in particular, user-centric systems, by using a single-user dataset?
@@
 The user data for this work has been extracted from the
-listening history of the corresponding author user in Last.fm
-\footnote[4]{https://www.last.fm/user/jimmydj2000}.
+listening history of the corresponding author user in Last.fm\footnote[5]{https://www.last.fm/user/jimmydj2000}.
 
@@
 we leverage the knowledge discovery potential
-of large historical listening logs, gathered from Last.fm.
+of large historical listening logs, gathered for the mentioned user from Last.fm.
 
@@
 
-Last.fm uses the term \emph{scrobble} to refer to a single track playback,
-at a particular moment.
-We queried the Last.fm API to download the user{'}s scrobble logs, reported from 2007 to 2022.
+Last.fm uses the term \emph{scrobble} to refer to the action of playing a track at a specific moment in time.
+Initially, Last.fm monitored user listening activity with a desktop application called \emph{Scrobbler}.
+Users used to install this application on their computers to monitor their activity on players such as Winamp or iTunes.
+With the advent of music streaming services, the possibilities for users to scrobble their music habits expanded.
+Integrations where developed to integrate the scrobbler into popular platforms,
+such as Spotify, YouTube, or SoundCloud.
+Mobile versions of the Scrobbler were developed for Android and iOS devices, and also open source initiatives flourished\footnote[6]{https://github.com/elamperti/OpenWebScrobbler}.
+
+For us, the first step to construct the dataset was to download the user listening activity.
+We queried the Last.fm API to download the user{'}s scrobbling logs, reported from 2007 to 2022.
 For each scrobble, we have gathered the following information:
@@
 \begin{itemize}
-\item Playback timestamp
-\item Track name
-\item Artist name
+\item Playback timestamp.
+\item Track name.
+\item Artist name.
 \item Track tags. If the track does not have any tags assigned,
-then artist tags have been used
+then artist tags were used.
 \end{itemize}
 
-For each tag-track mapping, Last.fm includes
+Last.fm maps each track (and artist) to a list of community-contributed tags.
+For each track-tag mapping, Last.fm includes
 a \emph{count} value, which indicates the popularity of the given tag for the track.
@@
 count value of 100.
-For example, if \emph{piano} is the most popular tag for a track,
+For example, if \emph{jazz club} is the most popular tag for a track,
 then the track might be probably
-associated to the following tuple \verb|(piano, 100)|.
+associated to the following tuple \verb|(jazz club, 100)|.
 
-Users normally listen to their favorite tracks several times,
-so the amount of unique tracks listened is much smaller
+Users tipically listen to their favorite tracks several times,
+so the amount of unique tracks played is much smaller
 than the number of track plays. In this case, the amount of
-individual tracks listened is about 20,000, and the number of scrobblings is, approximately, 90,000.
+individual tracks listened is about 20,000 and the number of scrobblings is, approximately, 90,000.
+Therefore, the user has listened to each song, approximately, 4.5 times on average.
 
@@
 Table \ref{table:spotify-features}.
-The reader can find further details about each feature in the Spotify API documentation
-\footnote[5]{
+The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
       \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
@@
 
-The Spotify API failed to provide audio features for a portion of the tracks.
-These tracks were filtered out from our experiments.
-After filtering tracks that miss Last.fm tags or Spotify audio features,
-our dataset contains 14,009 samples.
+\subsection{Filtering Missing Values}
+
+After gathering the data, we identified that the Spotify API had failed to provide audio features for a portion of the tracks.
+Similarly, Last.fm also returned no tags for another subset of the tracks.
+To prevent problems with missing values, we decided to filter out these tracks from the dataset, and therefore, from our experiments.
+After filtering tracks that were missing Last.fm tags or Spotify audio features,
+the dataset resulted in 14,009 samples.
+Compared to the original 20,000 unique tracks included in the listening history, approximately 6,000 songs were missing either Spotify or Last.fm data.
+
+\subsection{Dataset Comparison}
 
@@
 we explored the data to verify that the distribution of the Spotify audio features
-was comparable to other Spotify datasets.
+was comparable to larger, and possibly more balanced, Spotify datasets.
 In particular, we verified that the distribution of the features,
@@
 was comparable to the distribution of the \emph{Spotify Audio Features} Kaggle dataset.
-This dataset contains more tha 116,000 unique tracks, and includes audio features for each track.
-Our dataset present similar $\mu$ and $\sigma$ values as the Spotify Audio Features Kaggle dataset,
-as illustrated in Table \ref{table:audio-features-stats-kaggle} and Figure \ref{fig:audio-features-distribution-kaggle}.
-\footnote[6]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
+Our dataset presents similar $\mu$ and $\sigma$ values to the Spotify Audio Features Kaggle dataset,
+as illustrated in Table \ref{table:audio-features-stats-kaggle}
+and Figure \ref{fig:audio-features-distribution-kaggle}.
 
@@
 
-\subsection{Filtering Missing Values}
-
-Not every track of the approximately unique tracks 20,000 of the Last.fm user{'} listening history
-included Last.fm tags and Spotify audio features.
-Tracks without Last.fm tags or without Spotify audio features were filtered out.
-As a result, of the originally 20,000 tracks included in the listening history,
-our dataset final size resulted in 14,009 samples.
-
 \section{Experiments}
 
-We trained commonly used machine learning models to predict an audio feature, given the set of tags for a particular track.
+We trained commonly used machine learning models to predict Spotify audio features, given the set of Last.fm tags for a particular track.
 
@@
 
-The number of columns is limited to the top-K tags.
 Counting the total amount of Last.fm tags in the user collection resulted, initially, in more than five million tags.
@@
 
-The reduction algorithm is simple: calculate a weighted sum of all the tags appearances and pick the top 1000.
-The sum is weighted be- cause we use the count attribute.
-This attribute is present in every Last.fm track-tag association and provides a measure of the strength of a particular tag in a specific track.
-Note that this reduction is an initial approach, which, similar to other phases, can be extended or improved in the future.
-For this particular case, dimensionality reduction algorithms, such as PCA, are good candidates for forthcoming iterations of this work.
+The reduction algorithm uses a basic data aggregation algorightm: group the data by tag, aggregate by summing the \emph{count} values for each tag,
+order by the aggregated count, and finally select the top-$K$ items of the ranking.
 
-The input data passed is formatted in tabular format, as follows:
+We generated three versions of the dataset, with different values of $K$: 100, 1,000 and 10,000.
+
+We consider this reduction approach an initial approach.
+For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.
+
+After selecting the top-$K$ tags, the input data passed was formatted in tabular format, as follows:
 
 \begin{itemize}
-      \item Given that $Tags_{k}$ is the set of most $k$ frequent Last.fm tags in the user listening history
-            and, where each $tag \in Tags_{k}$.
+      \item Given that $Tags_{K}$ is the set of most $K$ frequent Last.fm tags in the user listening history
+            and, where each $tag \in Tags_{K}$.
       \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
@@
 \begin{itemize}
-      \item By including tag popularity: `rock 2, pop 1`.
-      \item By repeating tags based on popularity: `rock rock, pop`.
-      \item By ordering by popularity: `rock, pop`.
+      \item Ordering by count: \verb|"rock, pop"|.
+      \item Including tag count: \verb|"rock 2, pop 1"|.
+      \item Duplicating each tag $count$ times: \verb|"rock rock, pop"|.
 \end{itemize}
@@
 In this particular case, the $X$ values of the tabular input data are tokens.
-These tokens are obtained from passing the a string of concatenated Last.fm tags through a tokenizer.
-The formal definition of this data format is as follows:
+These tokens are obtained from passing the string of concatenated Last.fm tags through the GPT-2 tokenizer,
+as the following procedure explains:
 
 \begin{itemize}
-      \item Given that $X_l$ is the token vocabulary, where $l$ is the maximum vocabulary length.
+      \item Given that $X_L$ is the token vocabulary, where $L$ is the maximum vocabulary length.
       \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
@@
 
-\subsection{Text}
+Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
+
+The two tabular formats, with tags and tokens, were tested on the Naive Bayes and the Boosted tree regressors.
+
+\subsection{Text Strings}
 
 When using transformer models, the input data is a string.
-We must represent the Last.fm tags, which are initially in the $(tag name, tag popularity)$ form, to a a string.
+We must represent the Last.fm tags, which are defined as $(tag, count)$ tuples, as strings.
+To this end, we applied the same three transformations used in the tabular tokens formats,
+concatenate the tags by ordering by count, by including the tag count, and by duplicating the tags.
 
@@
       \begin{minipage}{\textwidth}
-      \caption{Text data format for tokens in XGBoost and Bayesian regressors}\label{text_format}%
+      \caption{Text data format for GPT-transformer. In this particular case, the tags have been concatenated by ordering by tag count}\label{text_format}%
       \begin{tabular}{@{}lllll@{}}
@@
 
-\begin{table}[h]
+\begin{table}[h!]
       \begin{center}
@@
 
+% INFO - Training Set: (8654, 14)
+% INFO - Validation Set: (2164, 14)
+% INFO - Test Set: (2705, 14)
+
+% TRAINING STATS
+% Num examples = 8654
+% Num Epochs = 10
+% Instantaneous batch size per device = 10
+% Total train batch size (w. parallel, distributed & accumulation) = 10
+% Gradient Accumulation steps = 1
+% Total optimization steps = 8660
+% Number of trainable parameters = 124441344
+
 \begin{table}[h!]
@@
       \midrule
-      GPT       & Duplicated\footnotemark[5]          & 0.157            & ...             & ...            & ...             & ...            \\
-      GPT       & Ordered\footnotemark[6]             & 0.149            & ...             & ...            & ...             & ...            \\
-      GPT       & Tags,Counts\footnotemark[7]         & \textbf{0.145}   & ...             & ...            & ...             & ...            \\
+      GPT       & Duplicated\footnotemark[5]          & 0.157            & 0.244           & 0.193          & 0.245           & 0.322          \\
+      GPT       & Ordered\footnotemark[6]             & 0.149            & \textbf{0.237}  & 0.188          & 0.235           & \textbf{0.297}          \\
+      GPT       & Tags,Counts\footnotemark[7]         & \textbf{0.145}   & \textbf{0.237}  & \textbf{0.187} & \textbf{0.233}  & 0.301          \\
 

### 2023-06-02 paper/main.tex
solved minor problems

@@
 Depending on the specific MIR application, researchers or practitioners expect different output values.
-For example, an application that extracts audio features might return values for the tempo, the key, or the sample rate.
+Applications that extract audio features typically return values related to the tempo, the key, or the sample rate, among others.
 In more complex applications, where, for example, machine learning is used,
@@
 
+The mapping between Last.fm and Spotify tracks is performed on an artist-track basis.
+The artist and track name extracted from Last.fm are used as parameters of the Spotify Search API.
+The Last.fm API provides a unique identifier, the \emph{MusicBrainz} ID.
+The Spotify API, however, does not provide this value.
+
 \subsection{Dataset Comparison}
@@
       \item Naive Bayes Regressor \cite{bayesian}
-      \item Fine-tuned GPT-2 model
+      \item Fine-tuned GPT-2 model \cite{radford2019language}
 \end{itemize}
@@
 was doable, but presented scalability problems.
+In addition to scalability limitations, classic machine learning models might not take advantage of using the full dataset.
+These models might even perform poorly if too many input features are provided.
+The reason for this is spurious relations or redundancy  between input features.
+Models might find relations that are not real, and latent, redundant variables might be accountable multiple times, which creates a bias in the output.
+Feature subject selection aims at solving these problems.
+
+% For example, a Bayesian model might provide low-quality results of provided with variables that are not conditionally independent.
+
 Therefore, we reduced the number of tags by picking a subset of the most relevant tags.
@@
 Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
-To tokenize tags, we have used the GTP2 tokenizer.
+To tokenize tags, we have used the GTP-2 tokenizer.
+
+Tokenizers are crucial elements in the preprocessing of text data.
+A tokenizer dissects a piece of text into smaller units, called tokens.
+These tokens can be words, subwords, or even syllables or characters.
+
+When breaking down the text into tokens, the tokenizer assigns a unique numerical identifier to each token.
+This IDs are based on the vocabulary that the tokenizer has been trained on.
+For example, when the tokenizer processes the \verb|"Hello world"| string, the Hello token might be assigned to ID \verb|123|
+and the world token might be assigned to ID \verb|34534|.
+The result of tokenizing \verb|Hello world| would be \verb|[123, 34534]|.
+
+Note that, although tokenizers are most commonly used in combination with transformer models, in this paper we test the possiblity of using theorem
+to preproccess data passed to Bayesian an Tree models.
+
 Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
@@
 
+Track mapping between Last.fm and Spotify is another opportunity for research and improvement.
+Although Last.fm provides the MusicBrainz unique identifier, but Spotify does not provide this value.
+The mapping was performed by using the track artist and name, but this approach resulted on about 30\% of the tracks not found on Spotify.
+
 \textcolor{red}{TODO}

### 2023-06-02 paper/main.tex
wip

@@
 processing, and knowledge discovery of information contained in music.
-MIR research intersects with other fields, such as computer science, signal processing, musicology, and sociology.
-The field covers applications in recommendation systems, music classification,
+MIR research covers a wide range of applications and intersects with other areas, such as computer science, signal processing, musicology, and sociology.
+Examples of MIR applications are recommendation systems, music classification,
 music source separation, and music generation, among others \cite{ramirez2020machine}.
@@
 MIR applications often attempt to extract information from the music audio signal,
-but analyzing associated metadata is also a common practice.
-Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based signal representations or hand-crafted audio features.
-Associated metadata, such as editorial information, lyrics, or user-generated content is usually in text format.
-However, metadata can be available as images of videos, for example, when analyzing album artwork, or music videos.
+although analyzing associated metadata is also a common practice.
+Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based signal representations (e.g. spectrograms),
+and sets of hand-crafted audio features, which are typically engineered by using domain knowledge (e.g. MFCC, rhythm, or tonal descriptors).
+
+The metadata associated to a music piece is available in multiple formats.
+Editorial information or lyrics, for example, are mostly available in text format.
+The ability to process images or videos is also required, for example, for analyzing album artwork, or music videos.
 
 Depending on the specific MIR application, researchers or practitioners expect different output values.
-Applications that extract audio features typically return values related to the tempo, the key, or the sample rate, among others.
-In more complex applications, where, for example, machine learning is used,
-applications might return the estimated emotion that a track produces on a listener, or the predicted music genre of a particular track.
+Applications that extract audio features typically return audio descriptors, namely values related to the tempo, the key, or the sample rate, among others.
+Open source libraries such as \emph{Librosa}\footnote[1]{https://librosa.org/} 
+and \emph{Essentia}\footnote[2]{https://essentia.upf.edu/index.html} offer methods to extract these values.
+Other applications might produce higher-level values, for example, by using machine learning techniques
+that estimate the emotion that a track induces, or the music genre of this track.
 
 Among potentially useful input and output values, research has proved Spotify audio features and Last.fm tags to be significant values to characterize music.
-Spotify audio features capture high-level information about the music signal.
-Examples of these values are energy, danceability, or valence, among others.
-Last.fm tags are text labels that users associate to particular tracks, artists or albums in the Last.fm website.
+Spotify audio features capture high-level information about the music signal,
+such as energy, danceability, or valence, among others.
+Last.fm tags are text labels that users associate to songs, artists, and albums via the Last.fm social platform.
 
-Both types of values have been used as input data mostly for classification tasks, such as music genre recognition,
-where, given Spotify features or Last.fm tasks, the model estimates the music genres of a particular track.
-Previous studies, however, have not used these values as target outputs, to the best of our knowledge.
-In particular, the focus of this article is on predicting Spotify audio features, given a set of Last.fm tags.
+Both Spotify audio features and Last.fm tags have been used as input data mostly for classification tasks, such as music genre recognition,
+where, given a set of Spotify audio features and/or Last.fm tags, the model estimates the music genre(s) of a particular track.
+Previous studies, however, have not experimented with these values as target outputs, to the best of our knowledge.
+This unexplored aspect reveals what we believe is a potential research opportunity in music analysis and recommendation.
 
-By predicting Spotify audio features, we explore the relationship between subjective perception and concrete musical features.
-This approach might help to identify patterns and hidden correlations between how music is percevied, consumed and discovered.
+In particular, this article focuses on predicting Spotify audio features, given a set of Last.fm tags.
+By predicting Spotify audio features, we explore the relationship between the subjective perception captured by Last.fm tags, and the concrete musical features that Spotify computes.
+This approach might help to identify patterns and hidden correlations between how music is perceived, consumed, and discovered.
 
-Additionally, predicted audio features could be used in recommendation systems to provide users with explainable recommendations.
-Music recommendations are tipically dificult to interpret from the perspective of the listener.
+Additionally, the predicted Spotify audio features could be used in recommendation systems to provide users with explainable recommendations.
+Music recommendations are typically difficult to interpret from the perspective of the listener.
 Users often get recommendations without meaningful explanations or justifications.
-Therefore, by using predicted Spotify features as basis for recommendation, we can explain users why the algorithm suggests a particular track.
+By predicting Spotify features as an intermediate step in the recommendation pipeline, we could use these features to explain users why the algorithm suggests a particular track.
 This process could be part of an explainable recommendation pipeline, where users enter a set of tags,
-and they receive the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
+and as a result they get the predicted audio features, the closest tracks to those features (as recommended tracks), and the distance values between each track and the predicted features.
 
@@
 
-\subsubsection{Last.fm Tags}
+\subsection{Last.fm Tags}
 
@@
 Last.fm tags have been a popular source of metadata for MIR tasks,
-because they potentionally contain subjective information related to the genre, mood, and style of music,
+because they potentially contain subjective information related to the genre, mood, and style of music,
 and might be used to characterize certain features of a music piece.
-Several studies have used Last.fm to predict music sentiment or mood, and to predict audio features.
+Additionally, Last.fm tags constitute a useful source of input knowledge when the audio signal is not available,
+for example, due to copyright limitations.
 
-Additionally, Last.fm tags constitute a useful source of input knownledge when the audio signal is not available,
-for example, due to copyright limitations.
+Several studies have used Last.fm to predict music sentiment, mood, and even audio features.
+For example, Laurier et al. analyze how Last.fm tags categorize mood.
+In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.
+
+% This study predicts Spotify features from Predictor Variables – Determinants of Music-Selection Behavior
+% https://www.static.tu.berlin/fileadmin/www/10002020/Dokumente/Abschlussarbeiten/Karnop_MasA.pdf
 
@@
 
-For example, Laurier et al. analyze how Last.fm tags categorize mood.
-In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.
-
 {\c{C}}ano and Morisio discuss how they create a dataset of music
@@
 In a similar direction, Bod{\'o} and Szil{\'a}gyi generate a dataset for lyrics genre classification
-by combining the Last.fm with MusicBrainz data \cite{bodo2018connecting}.
+by combining the Last.fm with \emph{MusicBrainz} data \cite{bodo2018connecting}.
 MusicBrainz is an online database of music editorial metadata\footnote[1]{https://musicbrainz.org/}.
@@
 
-\subsubsection{Spotify Audio Features}
+\subsection{Spotify Audio Features}
 

### 2023-06-02 paper/main.tex
wip

@@
 We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.
+\textcolor{red}{UNDER DEVELOPMENT}
+
+\section{Data Sources}
+
+\subsection{Last.fm Tags}
+
+Last.fm is an online music service for users to keep track of their music listening habits.
+Last.fm is also an online community where users tag artists, albums, and tracks, according their own taste and perception.
+
+Users apply these tags to
+categorize music from their own perspective, which
+means that tags do not fit into any structured ontology
+or data model.
+Tags can refer to any aspect that users consider as a valid descriptor, such as genre,
+emotion, or user listening context.
+
+For nearly two decades, users have been contributing to Last.fm by tagging tracks, albums, and artists with text labels.
+Although many of these descriptions are single-worded (e.g \emph{rock}, \emph{dance}, or \emph{happy}),
+users can also use short sentences to define a song, such as \emph{I like this track}, or \emph{on the beach}.
+
+\subsection{Spotify Audio Features}
+
+Spotify, one of the leaders in the music streaming industry, provides information about each track in their catalog via the Spotify developers API.
+One of the API endpoints is the audio features endpoint.
+
+The Spotify audio features are numerical values that represent high-level audio information computed from a specific
+track. These values characterize a track, musically speaking,
+by measuring musical aspects that, in many of these features, are related to the perspective of the user recommendation factors.
+For example, a danceability value of 0.95 means
+that a particular song is highly suitable for dancing.
+
+The features provided by the Spotify API are listed in
+Table \ref{table:spotify-features}.
+The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
+      \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
+}.
+
+\begin{table}[h!]
+      \centering
+      \caption{Spotify audio features. These features provide high-level musical information about a track.} \label{table:spotify-features}
+      \begin{tabular}{p{0.3\linewidth}p{0.6\linewidth}}
+          \toprule
+          \bfseries \textbf{Feature name} & \textbf{Description} \\
+          \midrule
+          \textbf{acousticness} & The track is acoustic. From 0 to 1 \\
+          \textbf{danceability} & The track encourages (or is adequate for) dancing. From 0 to 1 \\
+          \textbf{duration\_ms}  &  Duration in milliseconds \\
+          \textbf{energy}  &  The track is perceived as energetic. From 0 to 1\\
+          \textbf{instrumentalness}  &  The track is instrumental. From 0 to 1 \\
+          \textbf{key}  &  Key categories encoded as integers. From C (0) to 11 \\
+          \textbf{liveness}  &  The audience is audible. From 0 to 1\\
+          \textbf{loudness}  &  In decibels. From -60 to 0 \\
+          \textbf{mode}  & Major (1) or minor (0) \\
+          \textbf{speechiness}  & Does the track contain speeches? From 0 to 1 \\
+          \textbf{tempo}  & In beats per minute (BPM) \\
+          \textbf{valence} & How happy is the track (BPM). From 0 to 1 \\
+          \bottomrule
+      \end{tabular}
+  \end{table}
 
@@
 
-% Spotify, one of the leaders in the music streaming industry...
-
 The Spotify audio features have been used in multiple studies.
@@
 although these three Spotify features are relevant in terms of characterizing emotion, more features are needed for MER.
+% In reality, energy is not exactly arousal, but serves as its surrogate.
 
@@
 
-\subsection{Last.fm Tags}
-
-Last.fm is an online music service for users to keep track of their music listening habits.
-Last.fm is also an online community where users tag artists, albums, and tracks, according their own taste and perception.
-
-Users apply these tags to
-categorize music from their own perspective, which
-means that tags do not fit into any structured ontology
-or data model.
-Tags can refer to any aspect that users consider as a valid descriptor, such as genre,
-emotion, or user listenting context.
-
-For nearly two decades, users have been contributing to Last.fm by tagging tracks, albums, and artists with text labels.
-Although many of these descriptions are single-worded (e.g \emph{rock}, \emph{dance}, or \emph{happy}),
-users can also use short sentences to define a song, such as \emph{I like this track}, or \emph{on the beach}.
-
 \subsubsection{Last.fm Downloaded Data}
@@
 collected Spotify audio features for each one of those tracks.
-
-The Spotify audio features are numerical values that represent high-level audio information computed from a specific
-track. These values characterize a track, musically speaking,
-by measuring relevant musical aspects.
-For example, a danceability value of 0.95 means
-that a particular song is highly suitable for dancing.
-
-The features provided by the Spotify API are listed in
-Table \ref{table:spotify-features}.
-The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
-      \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
-}.
-
-\begin{table}[h!]
-      \centering
-      \caption{Spotify audio features. These features provide high-level musical information about a track.} \label{table:spotify-features}
-      \begin{tabular}{p{0.3\linewidth}p{0.6\linewidth}}
-          \toprule
-          \bfseries \textbf{Feature name} & \textbf{Description} \\
-          \midrule
-          \textbf{acousticness} & The track is acoustic. From 0 to 1 \\
-          \textbf{danceability} & The track encourages (or is adequate for) dancing. From 0 to 1 \\
-          \textbf{duration\_ms}  &  Duration in milliseconds \\
-          \textbf{energy}  &  The track is perceived as energetic. From 0 to 1\\
-          \textbf{instrumentalness}  &  The track is instrumental. From 0 to 1 \\
-          \textbf{key}  &  Key categories encoded as integers. From C (0) to 11 \\
-          \textbf{liveness}  &  The audience is audible. From 0 to 1\\
-          \textbf{loudness}  &  In decibels. From -60 to 0 \\
-          \textbf{mode}  & Major (1) or minor (0) \\
-          \textbf{speechiness}  & Does the track contain speeches? From 0 to 1 \\
-          \textbf{tempo}  & In beats per minute (BPM) \\
-          \textbf{valence} & How happy is the track (BPM). From 0 to 1 \\
-          \bottomrule
-      \end{tabular}
-  \end{table}
 

### 2023-06-03 paper/main.tex
improving sections 2 3 and 4

--- 2023-06-02 paper/main.tex
@@
 
+% TENSES: https://blog.wordvice.com/video-which-verb-tenses-should-i-use-in-a-research-paper/
+
 %%================================%%
@@
 We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.
+\textcolor{red}{UNDER DEVELOPMENT}
+
+\section{Data Sources}
+
+\subsection{Last.fm Tags}
+
+Last.fm is an online music community where users keep track of their music listening habits.
+Users apply tags to artists, tracks, and albums to categorize and describe music from their own perspective, which
+implies that the Last.fm tags space does not fit into any structured ontology or data model.
+A tag can refer to any aspect that users consider to be a valid descriptor, such as genre,
+emotion, or user listening context.
+
+For nearly two decades, many music aficionados have collectively contributed their own unique, personal interpretation, opinions and feelings
+as tags in Last.fm.
+Although many of these tags are single-worded descriptors (e.g \emph{rock}, \emph{dance}, or \emph{happy}),
+users also use short sentences to describe music, such as \emph{I like this track}, or \emph{on the beach}.
+
+Tags are available via the Last.fm API.
+
+\subsection{Spotify Audio Features}
+
+Spotify, one of the leaders in the music streaming industry, provides information about the tracks in their catalog via the Spotify developers API.
+Among the different data entities exposed, the API provides access to the \emph{Audio Features} for each track.
+
+The Spotify audio features are numerical values that represent high-level audio information computed from a specific
+track. These values characterize a track, musically speaking,
+by measuring musical aspects that, in many of these features, are related to the user perspective or recommendation factors.
+For example, a danceability value of 0.95 means
+that a particular song is highly suitable for dancing.
+
+The features provided by the Spotify API are listed in
+Table \ref{table:spotify-features}.
+While Spotify provides a description of the audio features.
+How they compute or estimate these values is not publicly available.
+The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
+      \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
+}.
+
+\begin{table}[h!]
+      \centering
+      \caption{Spotify audio features. These features provide high-level musical information about a track.} \label{table:spotify-features}
+      \begin{tabular}{p{0.3\linewidth}p{0.6\linewidth}}
+          \toprule
+          \bfseries \textbf{Feature name} & \textbf{Description} \\
+          \midrule
+          \textbf{acousticness} & The track is acoustic. From 0 to 1 \\
+          \textbf{danceability} & The track encourages (or is adequate for) dancing. From 0 to 1 \\
+          \textbf{duration\_ms}  &  Duration in milliseconds \\
+          \textbf{energy}  &  The track is perceived as energetic. From 0 to 1\\
+          \textbf{instrumentalness}  &  The track is instrumental. From 0 to 1 \\
+          \textbf{key}  &  Key categories encoded as integers. From C (0) to 11 \\
+          \textbf{liveness}  &  The audience is audible. From 0 to 1\\
+          \textbf{loudness}  &  In decibels. From -60 to 0 \\
+          \textbf{mode}  & Major (1) or minor (0) \\
+          \textbf{speechiness}  & Does the track contain speeches? From 0 to 1 \\
+          \textbf{tempo}  & In beats per minute (BPM) \\
+          \textbf{valence} & How happy is the track (BPM). From 0 to 1 \\
+          \bottomrule
+      \end{tabular}
+  \end{table}
 
 \section{Related Research}
+
+The objective of this study is to predict Spotify audio features from Last.fm tags, so the objective of this section is to
+explore existing literature that is related to the problem and review the facts that are important or relevant for our experimentation.
+
+The use of Last.fm tags and Spotify audio features has been common and prolific in studies
+that have applied machine learning to resolve MIR challenges.
+
+An important concept to consider is the arousal-valence cuadrants, present in music emotion recognition, 
+related to some studies that will be reviewed in the following sections, and also related to the valence and energy
+Spotify audio features.
 
@@
 Several studies have used Last.fm to predict music sentiment, mood, and even audio features.
-For example, Laurier et al. analyze how Last.fm tags categorize mood.
+For example, Laurier et al. analyzed how Last.fm tags categorize mood.
 In their study, they created a semantic mood space based on Last.fm tags \cite{laurier2009music}.
@@
 
-{\c{C}}ano and Morisio discuss how they create a dataset of music
+{\c{C}}ano and Morisio performed an analysis of Last.fm tags to create a dataset of music
 lyrics annotated with Last.fm tags.
-In the creation process, they conclude that Last.fm tags are mostly related to music genre
+In the creation process, they concluded that Last.fm tags are mostly related to music genre
 and positive moods \cite{ccano2017music}.
 
-In a similar direction, Bod{\'o} and Szil{\'a}gyi generate a dataset for lyrics genre classification
+In a similar direction, Bod{\'o} and Szil{\'a}gyi generated a dataset for lyrics genre classification
 by combining the Last.fm with \emph{MusicBrainz} data \cite{bodo2018connecting}.
@@
 
-So far, The \emph{Last.fm dataset} has been the most widely used Last.fm dataset\footnotemark[2] in research.
+Although different datasets that contain Last.fm tags have been created, so far,
+the \emph{Last.fm dataset}\footnotemark[2] has been the most widely used in research.
 This dataset is a complementary dataset of the Million Song Dataset (MSD) \cite{Bertin-Mahieux2011}.
@@
 
+\textcolor{red}{We introduced this section talking about machine learning but only refeferred to studies
+that use Last.fm tags in exploration and data set creation}
+
 \subsection{Spotify Audio Features}
 
-% Spotify, one of the leaders in the music streaming industry...
-
-The Spotify audio features have been used in multiple studies.
-Historically, these features were called the \emph{Echo Nest audio features}.
+Historically, the Spotify audio features features were called the \emph{Echo Nest audio features}.
 \emph{The Echo Nest} was an online music intelligence platform that provided users and clients with music analysis services.
@@
 Spotify acquired the Echo Nest in 2014.
-As a result, the Echo Nest API was shut down and Spotify made these audio features available via the Spotify API.
+As a result, the Echo Nest API was eventually deactivated and Spotify migrated these audio features to the Spotify API.
 
-Nowadays, published research is available under the two terms.
-Whereas the most recent studies refer to these audio features as the Spotify audio features,
+Nowadays, the two terms can be found in published research.
+Whereas the most recent studies refer to the Spotify audio features,
 earlier studies use the Echo Nest denomination.
@@
 
-While Spotify provides a description of the audio features, how they compute or estimate these values is not publicly available.
+Similarly to Last.fm, Spotify (or Echo Nest) audio features are commonly present in MIR research.
+For example, Wang and Horv{\'a}t used  these audio features to study the differences 
+between male and female artists \cite{wang2019gender}.
+\textcolor{red}{Y que consiguieron?}
 
-Wang and Horv{\'a}t use audio features to study differences between male and female artists \cite{wang2019gender}.
-
-Jamdar et al. used EchoNest audio features, combined with lyrics data to classify songs into emotion tags.
+Jamdar et al. used Echo Nest audio features, combined with lyrics data to classify songs into emotion tags.
 These classes were first defined based on a Last.fm tags emotion mapping \cite{jamdar2015emotion}.
@@
 They also achieve better performance on MER models by using their own top-100 features, and they determine that,
-although these three Spotify features are relevant in terms of characterizing emotion, more features are needed for MER.
+although these three Spotify features are relevant in terms of characterizing emotion,
+more features are needed to effectively recognize emotion.
+Another interesting observation of this study is the the identification of energy as a surrogate of arousal valence
 
+In general, Spotify audio features have been used as predictive input variables.
+We, to the best of our knowledge, are unaware of studies that use these features as target variables,
+or studies that have addressed the problem of audio features regression, based solely on Last.fm tags.
+
+Similarly to Last.fm tags, Spotify audio features can be found in a number of datasets.
 Publicly available datasets of Spotify audio features can be found online,
 as a result of open-source and research communities collecting data from the Spotify API and publishing the results.
-It is unclear, however, whether these published datasets violate the Spotify API terms of service.
+It is unclear, however, whether these published datasets completely meet the Spotify API terms of service.
 
@@
 
-Another example is the \emph{Spotify Audio Features} Kaggle dataset\footnote[4]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
+Another publicly available dataset is the \emph{Spotify Audio Features} Kaggle dataset\footnote[4]{https://www.kaggle.com/datasets/tomigelo/spotify-audio-features}.
 This dataset contains more than 116,000 unique tracks, and includes audio features for each track.
+
+For this reason, we chose to generate our own dataset.
 
@@
 % https://www.kaggle.com/datasets/andrewmvd/spotify-playlists
-
-In general, Spotify audio features have been used as predictive input variables.
-We, to the best of our knowledge, are unaware of studies that use these features as target variables,
-or studies that have addressed the problem of audio features regression, based solely on Last.fm tags.
 
@@
 
-% JAIME
-% I have posted a question on the Spotify Developer forum, relative to the release of a dataset
-% https://community.spotify.com/t5/Spotify-for-Developers/May-I-publish-an-open-source-dataset-that-contains-Spotify-Audio/td-p/5529259
+Before conducting experiments to predict audio features from tags,
+we constructed a dataset of Last.fm tags and Spotify audio features, indexed by track, by gathering the data from the Last.fm and Spotify APIs.
 
-Before conducting experiments to predict audio features from tags,
-we constructed a dataset, by gathering the data from the Last.fm and Spotify APIs.
+The tracks were selected from the listening history of a single user.
 
@@
 characterize the music preferences and listening context only for a single user.
-By training our system with single-user data, we also raise the following question: Is it possible to train
+Therefore, we extracted the data from the
+listening history of the corresponding author in Last.fm\footnote[5]{https://www.last.fm/user/jimmydj2000}.
+
+By training our system in a single-user space, we also raise the following question: Is it possible to train
 recommender systems, and in particular, user-centric systems, by using a single-user dataset?
+Additionally, we wanted to explore the idea of mimicking the fact that each human perceives music individually.
+If we train a system on data from different users, then the system would share the view of multiple individuals.
 
-The user data for this work has been extracted from the
-listening history of the corresponding author user in Last.fm\footnote[5]{https://www.last.fm/user/jimmydj2000}.
+Using a single-user data set might sound counterproductive in a machine learning scenario, specially considering
+how machine learning breakthroughs have attempted, and succeeded in many cases, to generalize in a particular problem.
+This is not objective of this study, which explores how a machine learning model can represent the music consumption experience of a single human.
+The model must be able to generalize, but only within the context of the user's musical taste, which be believe can be possible, given a sufficiently large listening history.
 
-Similar to other intelligent systems, recommender systems
-must be trained, by using user preference data, to produce
-suitable recommendations. For this study,
-we leverage the knowledge discovery potential
-of large historical listening logs, gathered for the mentioned user from Last.fm.
-
-Given the objective of our experiments,
-we created a dataset of Last.fm tags and Spotify audio features, indexed by track, by following these steps:
-
-\subsection{Last.fm Tags}
-
-Last.fm is an online music service for users to keep track of their music listening habits.
-Last.fm is also an online community where users tag artists, albums, and tracks, according their own taste and perception.
-
-Users apply these tags to
-categorize music from their own perspective, which
-means that tags do not fit into any structured ontology
-or data model.
-Tags can refer to any aspect that users consider as a valid descriptor, such as genre,
-emotion, or user listenting context.
-
-For nearly two decades, users have been contributing to Last.fm by tagging tracks, albums, and artists with text labels.
-Although many of these descriptions are single-worded (e.g \emph{rock}, \emph{dance}, or \emph{happy}),
-users can also use short sentences to define a song, such as \emph{I like this track}, or \emph{on the beach}.
-
-\subsubsection{Last.fm Downloaded Data}
+\subsection{Gathering Listening History and Tags from Last.fm}
 
 Last.fm uses the term \emph{scrobble} to refer to the action of playing a track at a specific moment in time.
-Initially, Last.fm monitored user listening activity with a desktop application called \emph{Scrobbler}.
+Last.fm started monitoring user listening activity with a desktop application called \emph{Scrobbler}.
 Users used to install this application on their computers to monitor their activity on players such as Winamp or iTunes.
@@
 such as Spotify, YouTube, or SoundCloud.
-Mobile versions of the Scrobbler were developed for Android and iOS devices, and also open source initiatives flourished\footnote[6]{https://github.com/elamperti/OpenWebScrobbler}.
+Mobile versions of the Scrobbler were also developed for Android and iOS devices, while open source initiatives flourished too\footnote[6]{https://github.com/elamperti/OpenWebScrobbler}.
 
@@
 \begin{itemize}
-\item Playback timestamp.
-\item Track name.
-\item Artist name.
-\item Track tags. If the track does not have any tags assigned,
-then artist tags were used.
+\item Playback timestamp
+\item Track name
+\item Artist name
+\item Track tags
 \end{itemize}
@@
 a \emph{count} value, which indicates the popularity of the given tag for the track.
-Last.fm normalizes this value in the 0-100 range, so the most popular tag for a track can have a
+Last.fm normalizes this value in the 0-100 range, so the most popular tag for a track is tipically associated with a
 count value of 100.
-For example, if \emph{jazz club} is the most popular tag for a track,
-then the track might be probably
-associated to the following tuple \verb|(jazz club, 100)|.
+For example, if \emph{jazz} is the most popular tag for a track,
+then the track might be probably associated to the following tuple \verb|(jazz, 100)|.
 
 Users tipically listen to their favorite tracks several times,
-so the amount of unique tracks played is much smaller
+so the amount of unique tracks played is smaller
 than the number of track plays. In this case, the amount of
-individual tracks listened is about 20,000 and the number of scrobblings is, approximately, 90,000.
-Therefore, the user has listened to each song, approximately, 4.5 times on average.
+individual tracks listened in a 15-year period is about \num{20000} and the number of scrobblings is, approximately, \num{90000}.
+Therefore, the user has listened to each song, approximately, \num{4.5} times on average.
 
-\subsection{Spotify Audio Features}
+\subsection{Gathering Spotify Audio Features}
 
-After gathering the listening history and track tags from Last.fm, and identifying the unique
-tracks that represent the user music collection, we
-collected Spotify audio features for each one of those tracks.
+After collecting the listening history and track tags from Last.fm, and identifying the unique
+tracks that represent the user music collection, we used the Spotify API to
+collect audio features for each one of the \num{20000} individual tracks.
 
-The Spotify audio features are numerical values that represent high-level audio information computed from a specific
-track. These values characterize a track, musically speaking,
-by measuring relevant musical aspects.
-For example, a danceability value of 0.95 means
-that a particular song is highly suitable for dancing.
-
-The features provided by the Spotify API are listed in
-Table \ref{table:spotify-features}.
-The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
-      \url{https://developer.spotify.com/documentation/web-api/reference/\#/operations/get-audio-features}
-}.
-
-\begin{table}[h!]
-      \centering
-      \caption{Spotify audio features. These features provide high-level musical information about a track.} \label{table:spotify-features}
-      \begin{tabular}{p{0.3\linewidth}p{0.6\linewidth}}
-          \toprule
-          \bfseries \textbf{Feature name} & \textbf{Description} \\
-          \midrule
-          \textbf{acousticness} & The track is acoustic. From 0 to 1 \\
-          \textbf{danceability} & The track encourages (or is adequate for) dancing. From 0 to 1 \\
-          \textbf{duration\_ms}  &  Duration in milliseconds \\
-          \textbf{energy}  &  The track is perceived as energetic. From 0 to 1\\
-          \textbf{instrumentalness}  &  The track is instrumental. From 0 to 1 \\
-          \textbf{key}  &  Key categories encoded as integers. From C (0) to 11 \\
-          \textbf{liveness}  &  The audience is audible. From 0 to 1\\
-          \textbf{loudness}  &  In decibels. From -60 to 0 \\
-          \textbf{mode}  & Major (1) or minor (0) \\
-          \textbf{speechiness}  & Does the track contain speeches? From 0 to 1 \\
-          \textbf{tempo}  & In beats per minute (BPM) \\
-          \textbf{valence} & How happy is the track (BPM). From 0 to 1 \\
-          \bottomrule
-      \end{tabular}
-  \end{table}
+The mapping between Last.fm and Spotify tracks was performed on an artist-track basis.
+For each track, the artist and track name extracted from Last.fm were used as parameters of the Spotify Search API.
+The Last.fm API provides a unique identifier, the \emph{MusicBrainz} ID.
+The Spotify API, however, does not provide this value.
 
@@
 
-After gathering the data, we identified that the Spotify API had failed to provide audio features for a portion of the tracks.
+After retrieving audio features, we identified that the Spotify API had failed to provide audio features for a portion of the tracks.
 Similarly, Last.fm also returned no tags for another subset of the tracks.
-To prevent problems with missing values, we decided to filter out these tracks from the dataset, and therefore, from our experiments.
+To prevent problems with missing values, we decided to filter out these tracks from the dataset.
 After filtering tracks that were missing Last.fm tags or Spotify audio features,
-the dataset resulted in 14,009 samples.
-Compared to the original 20,000 unique tracks included in the listening history, approximately 6,000 songs were missing either Spotify or Last.fm data.
-
-The mapping between Last.fm and Spotify tracks is performed on an artist-track basis.
-The artist and track name extracted from Last.fm are used as parameters of the Spotify Search API.
-The Last.fm API provides a unique identifier, the \emph{MusicBrainz} ID.
-The Spotify API, however, does not provide this value.
+the dataset resulted in \num{14009} samples.
+Compared to the original \num{20000} unique tracks included in the listening history, approximately \num{6000} songs were missing either Spotify or Last.fm data.
+In other words, about 70\% of the tracks in the user listening history included relevant information for the study.
 
@@
 described in Table \ref{table:audio-features-stats} and Figure \ref{fig:audio-features-distribution},
-was comparable to the distribution of the \emph{Spotify Audio Features} Kaggle dataset.
-Our dataset presents similar $\mu$ and $\sigma$ values to the Spotify Audio Features Kaggle dataset,
+was comparable to the distribution of the Spotify Audio Features Kaggle dataset.
+Our dataset, which we call \emph{Jimmydj2000 Single-user} dataset, presents mean ($\mu$) and standard deviation($\sigma$) values that are comparable to the same values of the Spotify Audio Features Kaggle dataset,
 as illustrated in Table \ref{table:audio-features-stats-kaggle}
@@
 
-\begin{table}[h!]
-      \begin{center}
-      \begin{minipage}{\textwidth}
-      \caption{Audio features description}\label{table:audio-features-stats}%
+\begin{table}[!ht]
+      \begin{minipage}{.5\linewidth}
+      \caption{Audio features description of the Jimmydj2000 Single-user dataset.}\label{table:audio-features-stats}
+            \centering
                   \begin{tabular}{@{}lll@{}}
@@
       \end{minipage}
-      \end{center}
-\end{table}
-\begin{table}[h!]
-      \begin{center}
-      \begin{minipage}{\textwidth}
-      \caption{Audio features description of \emph{Spotify Audio Features} Kaggle dataset}\label{table:audio-features-stats-kaggle}%
+      \begin{minipage}{.5\linewidth}
+      \caption{Audio features description of Spotify Audio Features Kaggle dataset.}\label{table:audio-features-stats-kaggle}%
+            \centering
                   \begin{tabular}{@{}lll@{}}
@@
       \end{minipage}
-      \end{center}
+
 \end{table}
+% \begin{table}[h!]
+%       \begin{center}
+
+%       \end{center}
+% \end{table}
 
@@
       \includegraphics[width=\textwidth]{images/features-distribution.png}
-      \caption{Distribution of audio features in our single-user dataset.}
+      \caption{Distribution of audio features in the Jimmydj2000 Single-user dataset.}
       \label{fig:audio-features-distribution}
@@
       \includegraphics[width=\textwidth]{images/feature-distribution-kaggle.png}
-      \caption{Distribution of audio features in the \emph{Spotify Audio Features} Kaggle dataset.}
+      \caption{Distribution of audio features in the Spotify Audio Features Kaggle dataset.}
       \label{fig:audio-features-distribution-kaggle}
@@
 
-We trained commonly used machine learning models to predict Spotify audio features, given the set of Last.fm tags for a particular track.
+We trained three commonly used machine learning models to predict Spotify audio features from Last.fm tags by using the
 

### 2023-06-03 paper/main.tex
add models description

@@
 was comparable to the distribution of the Spotify Audio Features Kaggle dataset.
-Our dataset, which we call \emph{Jimmydj2000 Single-user} dataset, presents mean ($\mu$) and standard deviation($\sigma$) values that are comparable to the same values of the Spotify Audio Features Kaggle dataset,
+Our dataset, which we call \emph{Last.fm Single-user} dataset, presents mean ($\mu$) and standard deviation($\sigma$) values that are comparable to the same values of the Spotify Audio Features Kaggle dataset,
 as illustrated in Table \ref{table:audio-features-stats-kaggle}
@@
       \begin{minipage}{.5\linewidth}
-      \caption{Audio features description of the Jimmydj2000 Single-user dataset.}\label{table:audio-features-stats}
+      \caption{Audio features description of the Last.fm Single-user dataset.}\label{table:audio-features-stats}
             \centering
@@
       \includegraphics[width=\textwidth]{images/features-distribution.png}
-      \caption{Distribution of audio features in the Jimmydj2000 Single-user dataset.}
+      \caption{Distribution of audio features in the Last.fm Single-user dataset.}
       \label{fig:audio-features-distribution}
@@
 
-We trained three commonly used machine learning models to predict Spotify audio features from Last.fm tags by using the
+We trained three commonly used machine learning models to predict Spotify audio features from Last.fm tags by using the Last.fm Single-user dataset.
+Considering that predicting the audio feature values is a regression problem, we tested the following models:
 
@@
       \item Boosted tree regressor \cite{xgboost}
-      \item Naive Bayes Regressor \cite{bayesian}
-      \item Fine-tuned GPT-2 model \cite{radford2019language}
+      \item Bayesian Ridge Regressor \cite{bayesian}
+      \item GPT-2 model (fine-tuning) \cite{radford2019language}
 \end{itemize}
+
+The \emph{Boosted tree regressor} is a specific implementation of the \emph{XGBoost} algorithm for regression tasks.
+This algorithm is an ensemble learning technique that combines simple multiple decision trees to
+create a stronger predictive model.
+This model has proved to be a powerful solution for classification and regression problems on high-dimensional, structured data.
+\textcolor{red}{REF}
+
+The \emph{Bayesian Ridge regressor} is a regression model that combines the principles of Bayesian statistics with ridge regression.
+Compared to linear regression, which assumes that the model estimated parameters have a deterministic value, the Bayesian Ridge regressor treats the model coefficients as variables with a prior distribution.
+By incorporating prior distributions into the learning process, the model is able to capture uncertainty.
+The ridge regression technique enables the model to handle noisy, collinear, structured data.
+
+The \emph{GTP-2} model is a transformer model.
+Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
+such as text generation and question answering.
+GPT-2 has been trained on a large corpus of text, and typically used as a language model to generate text that resembles human-written language.
+However, in this study we have experimented with GPT-2 as a regressor.
+Rather than retraining the model from scratch, we have fine tuned the model to behave as a regressor and  predict Spotify audio features.
+The basic idea behind this approach is to feed a string of concatenated Last.fm tags as input to the GPT-2 model, and then use the model's output as the predicted value of a specific Spotify audio feature.
+
+\subsection{Last.fm Tags Input Format}
 
@@
 Therefore, we tested different input formats.
-
-\subsection{Last.fm Tags Input Format}
 
@@
 
-\subsection{Boosted Tree Regressor}
+% \subsection{Boosted Tree Regressor}
 
-We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.
+% We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.
 
-\begin{table}[h!]
-      \begin{center}
-      \begin{minipage}{174pt}
-      \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
-      \begin{tabular}{@{}llll@{}}
-      \toprule
-      Parameter               & Value \\
-      \midrule
-      objective               & reg:squarederror  \\
-      base score              & 0.5 \\
-      booster                 & gbtree  \\
-      colsample bylevel       & 1 \\
-      colsample bynode        & 1 \\
-      colsample bytree        & 1 \\
-      gamma\footnotemark[1]   & 0 \\
-      learning rate           & 0.300000012 \\
-      max delta step          & 0 \\
-      max depth               & 6 \\
-      min child weight        & 1 \\
-      estimators              & 200  \\
-      n jobs                  & 12  \\
-      num parallel tree       & 1 \\
-      predictor               & auto  \\
-      random state            & 0 \\
-      reg alpha               & 0 \\
-      reg lambda              & 1 \\
-      scale pos weight        & 1 \\
-      subsample               & 2 \\
-      tree method             & auto  \\
-      \botrule
-      \end{tabular}
-      \footnotetext[1]{Minimum loss reduction required to make a further partition on a leaf node of the tree.}
-      \end{minipage}
-      \end{center}
-\end{table}
+% \begin{table}[h!]
+%       \begin{center}
+%       \begin{minipage}{174pt}
+%       \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
+%       \begin{tabular}{@{}llll@{}}
+%       \toprule
+%       Parameter               & Value \\
+%       \midrule
+%       objective               & reg:squarederror  \\
+%       base score              & 0.5 \\
+%       booster                 & gbtree  \\
+%       colsample bylevel       & 1 \\
+%       colsample bynode        & 1 \\
+%       colsample bytree        & 1 \\
+%       gamma\footnotemark[1]   & 0 \\
+%       learning rate           & 0.300000012 \\
+%       max delta step          & 0 \\
+%       max depth               & 6 \\
+%       min child weight        & 1 \\
+%       estimators              & 200  \\
+%       n jobs                  & 12  \\
+%       num parallel tree       & 1 \\
+%       predictor               & auto  \\
+%       random state            & 0 \\
+%       reg alpha               & 0 \\
+%       reg lambda              & 1 \\
+%       scale pos weight        & 1 \\
+%       subsample               & 2 \\
+%       tree method             & auto  \\
+%       \botrule
+%       \end{tabular}
+%       \footnotetext[1]{Minimum loss reduction required to make a further partition on a leaf node of the tree.}
+%       \end{minipage}
+%       \end{center}
+% \end{table}
 
-\subsection{Naive Bayes Regressor}
+% \subsection{Naive Bayes Regressor}
 
-The Naive Bayes Regressor, and in particular, Bayesian Ridge, is the model used for regression in this case.
+% The Naive Bayes Regressor, and in particular, Bayesian Ridge, is the model used for regression in this case.
 
-% COMMENT copied FROM BayesianRidge source code
-% https://github.com/scikit-learn/scikit-learn/blob/9aaed4987/sklearn/linear_model/_bayes.py#L24
-%
-%     There exist several strategies to perform Bayesian ridge regression. This
-%     implementation is based on the algorithm described in Appendix A of
-%     (Tipping, 2001) where updates of the regularization parameters are done as
-%     suggested in (MacKay, 1992). Note that according to A New
-%     View of Automatic Relevance Determination (Wipf and Nagarajan, 2008) these
-%     update rules do not guarantee that the marginal likelihood is increasing
-%     between two consecutive iterations of the optimization.
+% % COMMENT copied FROM BayesianRidge source code
+% % https://github.com/scikit-learn/scikit-learn/blob/9aaed4987/sklearn/linear_model/_bayes.py#L24
+% %
+% %     There exist several strategies to perform Bayesian ridge regression. This
+% %     implementation is based on the algorithm described in Appendix A of
+% %     (Tipping, 2001) where updates of the regularization parameters are done as
+% %     suggested in (MacKay, 1992). Note that according to A New
+% %     View of Automatic Relevance Determination (Wipf and Nagarajan, 2008) these
+% %     update rules do not guarantee that the marginal likelihood is increasing
+% %     between two consecutive iterations of the optimization.
 
-The training parameters are listed in table \ref{table_bayesian_training_params}.
+% The training parameters are listed in table \ref{table_bayesian_training_params}.
 
-\begin{table}[h]
-      \begin{center}
-      \begin{minipage}{174pt}
-      \caption{Training parameters for XGBoost regressor}\label{table_bayesian_training_params}%
-      \begin{tabular}{@{}llll@{}}
-      \toprule
-      Parameter                     & Value \\
-      \midrule
-      Maximum iterations            & 300   \\
-      Tolerance\footnotemark[1]     & \num{1e-03} \\
-      alpha 1                       & \num{1e-06} \\
-      alpha 2                       & \num{1e-06} \\
-      lambda 1                      & \num{1e-06} \\
-      lambda 2                      & \num{1e-06} \\
-      \botrule
-      \end{tabular}
-      \footnotetext[1]{Tolerance for the stopping criteria.}
-      \end{minipage}
-      \end{center}
-\end{table}
-
-\subsection{Fine-tuned Transformer}
-
-Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
-such as text generation and question answering.
-They were first introduced in 2017 by paper [REF: "Attention Is All You Need"]
-and have become one of the most popular models due to their ability
-to handle long-range dependencies and process variable-length inputs.
-
-One popular variant of transformers is the Generative Pre-trained Transformer 2 (GPT-2),
-which was introduced by OpenAI in 2019.
-GPT-2 is a language model that has been trained on a large corpus of text,
-and can generate coherent and fluent text that resembles human-written language.
-GPT-2 has achieved remarkable results in a wide range of natural language processing tasks, including text generation, machine translation, and question answering.
-
-In this study, we have used GPT-2 as a regressor to predict Spotify audio features from Last.fm tags.
-The basic idea behind this approach is to feed a string of concatenated Last.fm tags as input to the GPT-2 model, and then use the model's output as the predicted value of a specific Spotify audio feature.
-By training the GPT-2 model on a large dataset of Last.fm tags and corresponding Spotify audio features,
-we aim to learn the complex relationships between these two types of data and to use this knowledge to evaluate the prediction accuracy of Spotify audio features based on Last.fm tags.
-
-To train the GPT-2 model as a regressor, we used a supervised learning approach.
-Specifically, we collected a large dataset of Last.fm tags and corresponding Spotify audio features, and used this dataset to train the GPT-2 model to predict Spotify audio features from Last.fm tags.
-We used a mean squared error loss function to optimize the model's performance during training, and we also employed techniques such as early stopping and learning rate scheduling to prevent overfitting and improve generalization performance.
-\textcolor{red}{Under development...}
+% \begin{table}[h]
+%       \begin{center}
+%       \begin{minipage}{174pt}
+%       \caption{Training parameters for XGBoost regressor}\label{table_bayesian_training_params}%
+%       \begin{tabular}{@{}llll@{}}
+%       \toprule
+%       Parameter                     & Value \\
+%       \midrule
+%       Maximum iterations            & 300   \\
+%       Tolerance\footnotemark[1]     & \num{1e-03} \\
+%       alpha 1                       & \num{1e-06} \\
+%       alpha 2                       & \num{1e-06} \\
+%       lambda 1                      & \num{1e-06} \\
+%       lambda 2                      & \num{1e-06} \\
+%       \botrule
+%       \end{tabular}
+%       \footnotetext[1]{Tolerance for the stopping criteria.}
+%       \end{minipage}
+%       \end{center}
+% \end{table}
 

### 2023-06-03 paper/main.tex
draft experiments results

@@
 The preceding models require specific input formats.
-Therefore, we tested different input formats.
+In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GTP-2 expects text strings as input.
 
-Each individual sample in the dataset corresponds to a unique track,
+Each individual sample in the Last.fm singe-user dataset corresponds to a unique track,
 and contains the list of Last.fm tag-count tuples (e.g. \verb|[(electronic, 100), (dance, 45), ...]|)
 and the values of Spotify audio features.
-Before experimenting with machine learning models, we prepared the data in a number of different formats,
-each one suitable for specific models.
+The Last.fm single-user dataset was converted to the formats described in the following sections, and then fed into the corresponding models.
 
-\subsubsection{Tabular}
-With this format, the Last.fm tags are represented as a table.
+\subsubsection{Last.fm Tags as Table Columns}
+The tabular format represents the Last.fm tags as columns in a table.
 Each tag is defined by a column and each cell contains the count value of a tag for a track.
-A cell is 0 if a tag is missing for a track.
+If a $tag$ is not present for a particular $track$, then cell $c_{track,tag}$ is \num{0}.
 
 Counting the total amount of Last.fm tags in the user collection resulted, initially, in more than five million tags.
-We quickly confirmed that building a tabular data set, in which every row contains millions of columns (Last.fm tags)
-was doable, but presented scalability problems.
-In addition to scalability limitations, classic machine learning models might not take advantage of using the full dataset.
+Under this high-dimensionality scenario, building a tabular data set, in which every row contains millions of columns (i.e.
+Last.fm tags) was theoretically possible, but presented scalability problems.
+In addition to scalability limitations, classic machine learning models might not take advantage of using the full collection of tags present in the user history.
 These models might even perform poorly if too many input features are provided.
 The reason for this is spurious relations or redundancy  between input features.
-Models might find relations that are not real, and latent, redundant variables might be accountable multiple times, which creates a bias in the output.
+Models might find relations that are not real, and latent, redundant variables might be accountable multiple times, which can lead to a biased outputs.
 Feature subject selection aims at solving these problems.
 
-% For example, a Bayesian model might provide low-quality results of provided with variables that are not conditionally independent.
-
 Therefore, we reduced the number of tags by picking a subset of the most relevant tags.
-
-The reduction algorithm uses a basic data aggregation algorightm: group the data by tag, aggregate by summing the \emph{count} values for each tag,
-order by the aggregated count, and finally select the top-$K$ items of the ranking.
-
-We generated three versions of the dataset, with different values of $K$: 100, 1,000 and 10,000.
-
-We consider this reduction approach an initial approach.
+We used a the feature selection algorithm by using a basic data aggregation algorithm: grouping the data by tag, aggregating by summing the \emph{count} values for each tag,
+ordering by the aggregated count, and finally selecting the top-$K$ tags of the ranking.
+Three values of $K$, were used, thus generating three subsets: \num{100}, \num{1000}, and \num{10000}.
+We consider this reduction approach an initial approach approach in our experiments.
 For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.
 
-After selecting the top-$K$ tags, the input data passed was formatted in tabular format, as follows:
+After selecting the top-$K$ tags, the Last.fm single-user dataset was formatted as follows:
 
@@
       \item Given that $Tags_{K}$ is the set of most $K$ frequent Last.fm tags in the user listening history
-            and, where each $tag \in Tags_{K}$.
+            , where each $tag \in Tags_{K}$.
       \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
@@
       \begin{itemize}
-            \item $X_{track},_{tag}$ is the strengh of $tag$ for $track$. This value is in the $0-100$ range.
+            \item $X_{tag,track}$ is the $count$ value of $tag$ for $track$. This value is in the $0-100$ range.
             \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
@@
 
-When generating training data by track, the tabular formats present sparsity problems.
+The main drawback of this representation is data sparsity.
+For most of the tracks, many columns are \num{0}.
 
-For tabular representations, we need to defined a fixed set of columns as tags.
-For most of tracks, most columns are 0.
+This format was tested with the Bayesian Ridge and the Boosted Tree regressors.
 
-The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
-(e.g., m * n for an m * n matrix) is called the sparsity of the matrix.
+% The sparsity of a matrix is the number of zero-valued elements divided by the total number of elements
+% (e.g., m * n for an m * n matrix) is called the sparsity of the matrix.
 
-\subsubsection{Tabular Tokens}
+\subsubsection{Tokens as Table Columns}
 
-Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
-To tokenize tags, we have used the GTP-2 tokenizer.
-
+This format is also structured, but the input data is a set of tokens instead of tag count values.
+These \emph{tokens} are the result of passing the tags, concatenated as a string, through a tokenizer.
 Tokenizers are crucial elements in the preprocessing of text data.
 A tokenizer dissects a piece of text into smaller units, called tokens.
-These tokens can be words, subwords, or even syllables or characters.
+These tokens can be words, subwords, or even characters.
+
+% Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
+% To tokenize tags, we have used the GTP-2 tokenizer.
 
 When breaking down the text into tokens, the tokenizer assigns a unique numerical identifier to each token.
-This IDs are based on the vocabulary that the tokenizer has been trained on.
-For example, when the tokenizer processes the \verb|"Hello world"| string, the Hello token might be assigned to ID \verb|123|
-and the world token might be assigned to ID \verb|34534|.
-The result of tokenizing \verb|Hello world| would be \verb|[123, 34534]|.
+These identifiers are based on the vocabulary that the tokenizer has been trained on.
+For example, when the tokenizer processes the \verb|"pop rock"| string, the \verb|pop| token might be assigned to ID \verb|123|
+and the \verb|rock| token might be assigned to ID \verb|34534|.
+Consequently, the result of tokenizing \verb|pop rock| would be \verb|[123, 34534]|.
 
-Note that, although tokenizers are most commonly used in combination with transformer models, in this paper we test the possiblity of using theorem
-to preproccess data passed to Bayesian an Tree models.
+Note that, although tokenizers are most commonly used in combination with transformer models, in this paper we test the possibility of using a tokenizer to preproccess the data passed to the models that require structured data.
 
-Because the tokenizer requires a string as input, we have converted the set of tags for each track into a string.
-To \emph{stringify} the tags, we have concatenated tags with multiple strategies:
+Because the tokenizer requires a string as the input, we converted the set of tags for each track into a string, in a process that we called \emph{stringification}.
+To \emph{stringify} the tags, we concatenated Last.fm tags by following these three strategies:
 
 \begin{itemize}
-      \item Ordering by count: \verb|"rock, pop"|.
-      \item Including tag count: \verb|"rock 2, pop 1"|.
-      \item Duplicating each tag $count$ times: \verb|"rock rock, pop"|.
+      \item Order by count: \verb|"rock, pop"|.
+      \item Include tag count: \verb|"rock 2, pop 1"|.
+      \item Duplicate tags $count$ times: \verb|"rock rock, pop"|.
 \end{itemize}
 
-In this particular case, the $X$ values of the tabular input data are tokens.
+In this particular case, the $X$ values of the tabular input data are the token IDs.
 These tokens are obtained from passing the string of concatenated Last.fm tags through the GPT-2 tokenizer,
@@
 \begin{itemize}
+      \item Given that $S$ is the stringification strategy.
       \item Given that $X_L$ is the token vocabulary, where $L$ is the maximum vocabulary length.
       \item Given that $Audio$ is the set of Spotify audio features, where each $feat \in Audio$.
-      \item For each $track$:
+      \item For each $track$ and $S$:
       \begin{itemize}
-            \item $X_{track},_{n}$ is token found at position $n$, after tokenizing the tags string.
-            \item $y_{track},_{feature}$ is the value of the audio feature $y$ for $track$.
+            \item $tags_{track,str}$ is the string of concatenated tags produced by strategy $S$.
+            \item $tokens_{track}$ is the list of token IDs produced after tokenizing $tags_{str}$.
+            \item $X_{n,track}$ is token ID found at position $n$ of $tokens_{track}$.
+            \item $y_{feature, track}$ is the value of the audio feature $y$ for $track$.
       \end{itemize}
@@
 Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
+\textcolor{red}{Decir cual es el nombre de este limite (maxsequence length?)}
 
-The two tabular formats, with tags and tokens, were tested on the Naive Bayes and the Boosted tree regressors.
+Similar to the previous structured format, this format was also used in the Bayesian and Boosted Tree regressors.
+In contrast to the previous format, this format does not present sparsity.
+It incorporates however, other problems, such as the lack of normalization of the values, which do not represent scalars, but IDs.
+It is also harder to inspect, because the meaning of each token is unknown.
 
-\subsection{Text Strings}
+\subsubsection{Text Strings}
 
 When using transformer models, the input data is a string.
-We must represent the Last.fm tags, which are defined as $(tag, count)$ tuples, as strings.
+To use GPT-2, we had to represent the Last.fm tags, which are defined as $(tag, count)$ tuples, as strings.
 To this end, we applied the same three transformations used in the tabular tokens formats,
@@
 
+The experiments tested the three models, and the three values of $K$: \num{100}, \num{1000}, \num{10000}.
+In the Bayesian Ridge and Boosted tree regressors, we tested the two structured data formats: tags and tokens as columns.
+In the GTP-2 model, we tested the text data format.
+
+For the formats that required the concatenation of tags into text strings, we tested the three different stringification strategies.
+
+The data was split into a training set (\num{8654} samples), a validation set(\num{2164} samples), and a test set (\num{2705} samples).
+The quality of the model was evaluated by using the root mean squared error (RMSE).
+This metric was computed on the test set.
+
 Table \ref{table:experiment_results} summaries the experiment results.
@@
       Tree      & 1000 tags                           & 0.149            &\textbf{0.249}   & 0.184          & 0.236           & 0.291          \\
-      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  &\textbf{0.235}   &\textbf{0.290}  \\
+      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  & 0.235           &\textbf{0.290}  \\
       Tree      & 100 tokens D                        & 0.274            & 0.274           & 0.212          & 0.256           & 0.330          \\
-      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & 0.220           & 0.339          \\
+      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & \textbf{0.220}  & 0.339          \\
       Tree      & 10000 tokens D                      & 0.172            & 0.276           & 0.217          & 0.266           & 0.342          \\
@@
       Tree      & 100 tokens TC                       & 0.172            & 0.280           & 0.211          & 0.262           & 0.342          \\
-      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & 0.220           & 0.344          \\
+      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & \textbf{0.220}  & 0.344          \\
       Tree      & 10000 tokens TC                     & 0.175            & 0.281           & 0.214          & 0.270           & 0.345          \\
@@
 \end{table}
+
+The Spotify feature that the models had less trouble estimating was the danceability.
+The instrumentalness, however, was the feature that presented the highest deviations.
+The GTP-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
+The Boosted tree was the best model to predict the instrumentalness, with a RMSE value of \num{0.29}.
+
+In general, the three models performed similarly, with GPT-2 achieving the slightly better results in danceability and acousticness.
+The Boosted Tree regressor was the best model to predict energy, valence, and instrumentalness.
+
+The two regressors achieved the best results when using the \emph{Tags as columns} format, mostly with \num{10000} tags.
+Interestingly, the Boosted Tree regressor achieved better results with tokens when estimating valence.
+Even better results than GPT-2.
+In the rest of the features, the regressors worked better with tags as columns.
+
+Using a higher number of predictor variables was not a synonym of lower error values when using tokens a columns.
+
+The Bayesian regressor was a bit more sensitive to changes in tag formating.
+This model tended to work better with data formatted by using the Tag Count stringification method.
+The Boosted Tree regressor presented less sensitivity to format changes, although the model generally performed better with the Tags as columns format too.
+
+The Duplicated stringification method consistently performed worse with the GTP-2 transformer model.
 
@@
 
+We consider this reduction approach an initial approach approach in our experiments.
+For this particular aspect, dimensionality reduction algorithms, such as PCA, are good candidates for future work.
+
 \textcolor{red}{TODO}

### 2023-06-03 paper/main.tex
fixes
more fixes

@@
 \end{table}
+
+\textcolor{red}{what is the limit on the sequence length here?}
 
@@
       Bayes     & 100 tokens D\footnotemark[2]        & 0.307            & 0.307           & 0.238          & 0.281           & 0.383          \\
+      Bayes     & 1000 tokens D                       & 0.201            & 0.315           & 0.249          & 0.297           & 0.399          \\
       Bayes     & 10000 tokens D                      & 0.359            & 0.507           & 0.394          & 0.479           & 0.613          \\
-      Bayes     & 1000 tokens D                       & 0.201            & 0.315           & 0.249          & 0.248           & 0.399          \\
       Bayes     & 100 tokens O\footnotemark[3]        & 0.191            & 0.305           & 0.237          & 0.282           & 0.376          \\
@@
       Tree      & 100 tags                            & 0.154            & 0.257           & 0.188          & 0.240           & 0.302          \\
-      Tree      & 1000 tags                           & 0.149            &\textbf{0.249}   & 0.184          & 0.236           & 0.291          \\
-      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  & 0.235           &\textbf{0.290}  \\
+      Tree      & 1000 tags                           & 0.149            &\textbf{0.249}   & 0.184          & 0.236           & 0.292          \\
+      Tree      & 10000 tags                          &\textbf{0.148}    & 0.250           &\textbf{0.181}  & \textbf{0.235}  &\textbf{0.291}  \\
       Tree      & 100 tokens D                        & 0.274            & 0.274           & 0.212          & 0.256           & 0.330          \\
-      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & \textbf{0.220}  & 0.339          \\
+      Tree      & 1000 tokens D                       & 0.173            & 0.278           & 0.215          & 0.268           & 0.339          \\
       Tree      & 10000 tokens D                      & 0.172            & 0.276           & 0.217          & 0.266           & 0.342          \\
@@
       Tree      & 100 tokens TC                       & 0.172            & 0.280           & 0.211          & 0.262           & 0.342          \\
-      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & \textbf{0.220}  & 0.344          \\
+      Tree      & 1000 tokens TC                      & 0.174            & 0.282           & 0.215          & 0.268           & 0.344          \\
       Tree      & 10000 tokens TC                     & 0.175            & 0.281           & 0.214          & 0.270           & 0.345          \\
@@
 
-The Spotify feature that the models had less trouble estimating was the danceability.
+The Spotify feature that the models had less trouble estimating was the danceability, and the energy.
 The instrumentalness, however, was the feature that presented the highest deviations.
+Acousticness and valence also presented high RMSE values.
+
 The GTP-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
@@
 
-In general, the three models performed similarly, with GPT-2 achieving the slightly better results in danceability and acousticness.
-The Boosted Tree regressor was the best model to predict energy, valence, and instrumentalness.
+In general, the three models performed similarly, with GPT-2 achieving slightly better results in danceability, acousticness, and valence.
+The Boosted Tree regressor was the best model to predict energy, and instrumentalness.
 
@@
 
-The Bayesian regressor was a bit more sensitive to changes in tag formating.
+The Bayesian regressor was a bit more sensitive to changes in tag formatting.
 This model tended to work better with data formatted by using the Tag Count stringification method.

### 2023-06-19 paper/main.tex
apply suggestions from Philippe

@@
 between male and female artists \cite{wang2019gender}.
-\textcolor{red}{Y que consiguieron?}
+\textcolor{red}{And what did they achieve?}
 
@@
 
-For this reason, we chose to generate our own dataset.
-
+There are various reasons why we chose to generate our own dataset.
+First, we wanted to generate a dataset that combined both Last.fm tags and Spotify audio features.
+Second, there is lack of clarity regarding the conditions under which Spotify allows the use of the audio features.
+Third, we wanted to explore a research line where machine learning models are trained by using the listening history of a single user. 
 % Another dataset created from the Spotify API and published to Kaggle:
@@
 
-% Additionally, language models, such as GTP-2 have shown promising results in various
+% Additionally, language models, such as GPT-2 have shown promising results in various
 % natural language processing (NLP) and generation.
 
-% In this particular study, GTP-2 has been fine-tuned for regression tasks,
+% In this particular study, GPT-2 has been fine-tuned for regression tasks,
 % and has shown good performance in predicting audio features from metadata.
@@
 % This paper aims to apply the boosted tree regressor,
-%  Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
+%  Bayesian ridge regressor, and GPT-2 models to predict Spotify audio features from Last.fm tags.
 % The experiments compare the performance of these models and evaluate their effectiveness
@@
 
-Users tipically listen to their favorite tracks several times,
+Users typically listen to their favorite tracks several times,
 so the amount of unique tracks played is smaller
@@
 \end{table}
-% \begin{table}[h!]
-%       \begin{center}
-
-%       \end{center}
-% \end{table}
 
@@
 
+\subsection{Models}
+
 The \emph{Boosted tree regressor} is a specific implementation of the \emph{XGBoost} algorithm for regression tasks.
@@
 
-The \emph{GTP-2} model is a transformer model.
+The \emph{GPT-2} model is a transformer model.
 Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
@@
 
+\subsubsection{Models Parametrization}
+
+Due to the different models and input encodings that we employed in the models, we decided to limit the dimensionality of the experiments, 
+by using mostly the default hyperparameters that the most common libraries set for these models.
+
+For the Bayesian ridge regressor, we used the default hyperparameters set by the \emph{scikit-learn} library.
+The training parameters for the Bayesian ridge are listed in table \ref{table_bayesian_training_params}.
+
+To fine tune the GPT-2 model, we used the \emph{GPT2ForSequenceClassification} model of the \emph{Transformers} Python library.
+The parameters used are listed in table \ref{table_gpt2_training_params}.
+
+\begin{table}[!ht]
+      \begin{minipage}{.5\linewidth}
+            \caption{Training parameters for Bayesian Ridge regressor}\label{table_bayesian_training_params}%
+            \centering
+                  \begin{tabular}{@{}ll@{}}
+                  \toprule
+                  Parameter                     & Value \\
+                  \midrule
+                  Maximum iterations            & 300   \\
+                  Tolerance\footnotemark[1]     & \num{1e-03} \\
+                  alpha 1                       & \num{1e-06} \\
+                  alpha 2                       & \num{1e-06} \\
+                  lambda 1                      & \num{1e-06} \\
+                  lambda 2                      & \num{1e-06} \\
+                  \botrule
+                  \end{tabular}
+            \footnotetext[1]{Tolerance for the stopping criteria.}
+            \caption{Training parameters for GPT-2 regressor}\label{table_gpt2_training_params}%
+            \centering
+                  \begin{tabular}{@{}ll@{}}
+                  \toprule
+                  Parameter                     & Value \\
+                  \midrule
+                  Batch size                    & 10 \\
+                  Problem type                  & regression \\
+                  Epochs                        & 10 \\
+                  Sequence length (tokens)      & 256 \\
+                  Tokenizer                     & GPT-2 tokenizer \\
+                  \botrule
+                  \end{tabular}
+      \end{minipage}
+      \begin{minipage}{.5\linewidth}
+            \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
+            \centering
+                  \begin{tabular}{@{}ll@{}}
+                  \toprule
+                  Parameter               & Value \\
+                  \midrule
+                  objective               & reg:squarederror  \\
+                  base score              & 0.5 \\
+                  booster                 & gbtree  \\
+                  colsample bylevel       & 1 \\
+                  colsample bynode        & 1 \\
+                  colsample bytree        & 1 \\
+                  gamma\footnotemark[1]   & 0 \\
+                  learning rate           & 0.3 \\
+                  max delta step          & 0 \\
+                  max depth               & 6 \\
+                  min child weight        & 1 \\
+                  estimators              & 200  \\
+                  n jobs                  & 12  \\
+                  num parallel tree       & 1 \\
+                  predictor               & auto  \\
+                  random state            & 0 \\
+                  reg alpha               & 0 \\
+                  reg lambda              & 1 \\
+                  scale pos weight        & 1 \\
+                  subsample               & 2 \\
+                  tree method             & auto  \\
+                  \botrule
+                  \end{tabular}
+            \footnotetext[1]{Tolerance for the stopping criteria.}
+      \end{minipage}
+\end{table}
+
+Similarly, for the boosted tree regressor, we used the default parameters set in the \emph{xgboost} Python library.
+We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.
+
 \subsection{Last.fm Tags Input Format}
 
+\textcolor{red}{TODO: Create diagram}
+
 The preceding models require specific input formats.
-In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GTP-2 expects text strings as input.
+In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GPT-2 expects text strings as input.
 
@@
 % Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
-% To tokenize tags, we have used the GTP-2 tokenizer.
+% To tokenize tags, we have used the GPT-2 tokenizer.
 
@@
 Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
-\textcolor{red}{Decir cual es el nombre de este limite (maxsequence length?)}
+\textcolor{red}{Describe the name of this particular limit (maxsequence length?)}
 
@@
 % \subsection{Boosted Tree Regressor}
-
-% We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.
-
-% \begin{table}[h!]
-%       \begin{center}
-%       \begin{minipage}{174pt}
-%       \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
-%       \begin{tabular}{@{}llll@{}}
-%       \toprule
-%       Parameter               & Value \\
-%       \midrule
-%       objective               & reg:squarederror  \\
-%       base score              & 0.5 \\
-%       booster                 & gbtree  \\
-%       colsample bylevel       & 1 \\
-%       colsample bynode        & 1 \\
-%       colsample bytree        & 1 \\
-%       gamma\footnotemark[1]   & 0 \\
-%       learning rate           & 0.300000012 \\
-%       max delta step          & 0 \\
-%       max depth               & 6 \\
-%       min child weight        & 1 \\
-%       estimators              & 200  \\
-%       n jobs                  & 12  \\
-%       num parallel tree       & 1 \\
-%       predictor               & auto  \\
-%       random state            & 0 \\
-%       reg alpha               & 0 \\
-%       reg lambda              & 1 \\
-%       scale pos weight        & 1 \\
-%       subsample               & 2 \\
-%       tree method             & auto  \\
-%       \botrule
-%       \end{tabular}
-%       \footnotetext[1]{Minimum loss reduction required to make a further partition on a leaf node of the tree.}
-%       \end{minipage}
-%       \end{center}
-% \end{table}
 
@@
 
-% The training parameters are listed in table \ref{table_bayesian_training_params}.
-
-% \begin{table}[h]
-%       \begin{center}
-%       \begin{minipage}{174pt}
-%       \caption{Training parameters for XGBoost regressor}\label{table_bayesian_training_params}%
-%       \begin{tabular}{@{}llll@{}}
-%       \toprule
-%       Parameter                     & Value \\
-%       \midrule
-%       Maximum iterations            & 300   \\
-%       Tolerance\footnotemark[1]     & \num{1e-03} \\
-%       alpha 1                       & \num{1e-06} \\
-%       alpha 2                       & \num{1e-06} \\
-%       lambda 1                      & \num{1e-06} \\
-%       lambda 2                      & \num{1e-06} \\
-%       \botrule
-%       \end{tabular}
-%       \footnotetext[1]{Tolerance for the stopping criteria.}
-%       \end{minipage}
-%       \end{center}
-% \end{table}
-
 \subsection{Experiments Execution and Results}
@@
 In the Bayesian Ridge and Boosted tree regressors, we tested the two structured data formats: tags and tokens as columns.
-In the GTP-2 model, we tested the text data format.
+In the GPT-2 model, we tested the text data format.
 
@@
 
+\textcolor{red}{TODO: as you have std deviations for your RMSE, can you say that
+.291 is really better than .297 ? (idem for other results)}
+
 The Spotify feature that the models had less trouble estimating was the danceability, and the energy.
@@
 
-The GTP-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
+The GPT-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
 The Boosted tree was the best model to predict the instrumentalness, with a RMSE value of \num{0.29}.
@@
 
-The Duplicated stringification method consistently performed worse with the GTP-2 transformer model.
+The Duplicated stringification method consistently performed worse with the GPT-2 transformer model.
 

### 2023-07-01 paper/main.tex
add abstract and piece of conclusions

@@
 
-\affil[1]{\orgdiv{Departamento de Sistemas Informáticos}, \orgname{Universidad de Castilla-La Mancha}, \orgaddress{\street{Campus universitario s/n}, \city{Albacete}, \postcode{02071}, \country{Spain}}}
-
 %%==================================%%
@@
 \abstract{
-    In this paper, we discuss a number of experiments to analyze the
-    suitability of music label representations to predict certain audio features,
-    such as danceability, loudness, or acousticness \dots
+      Music information retrieval (MIR) is an interdisciplinary research field that focuses on the extraction, processing,
+      and knowledge discovery of information contained in music.
+      While previous studies have utilized Spotify audio features
+      and Last.fm tags as input values for classification tasks, such as music genre recognition,
+      their potential as target values has remained unexplored.
+      In this article, we address this notable gap in the research landscape by proposing a novel approach to predict
+      Spotify audio features based on a set of Last.fm tags.
+      By predicting audio features, we aim to explore the relationship between subjective perception and concrete musical features,
+      shedding light on patterns and hidden correlations between how music is perceived, consumed, and discovered.
+      Additionally, the predicted audio features can be leveraged in recommendation systems
+      to provide users with explainable recommendations, bridging the gap between algorithmic suggestions and user understanding.
+      Our experiments involve training models such as GPT-2, XGBRegressor, and Bayesian Ridge regressor to predict
+      Spotify audio features from Last.fm tags.
+      Through our findings, we contribute to the advancement of MIR research by demonstrating the potential of Last.fm
+      tags as target values and paving the way for future research on the connection between subjective and objective music characterization.
+      Our approach holds promise for both listeners and researchers,
+      offering new insights into the intricate relationship between perception and audio signal in music.
 }
@@
       \begin{minipage}{\textwidth}
-      \caption{Experiment results. Cells values correspond to the RMSE value.}\label{table:experiment_results}%
+      \caption{Experiment results. Cells values correspond to the RMSE value.
+      Highlighted values correspond to the best RMSE value achieved by each model, for each variable.}
+      \label{table:experiment_results}%
       \begin{tabular}{@{}lllllll@{}}
@@
 
-\textcolor{red}{TODO: as you have std deviations for your RMSE, can you say that
-.291 is really better than .297 ? (idem for other results)}
-
 The Spotify feature that the models had less trouble estimating was the danceability, and the energy.
@@
 The mapping was performed by using the track artist and name, but this approach resulted on about 30\% of the tracks not found on Spotify.
+
+The standard deviation values that are displayed in figure \ref{fig:rmse_by_feature_and_input} are in the 0.1-0.2 margin.
+Considering this factor, we argue that the RMSE values that some models manifest are, in fact, noteworthy.
+For example, the best value for instrumentalness is \num{0.291}.
+Considering that the mean and standard deviation for all the experiments with this variable are, respectively,\num{0.414} and\num{0.106}, the z-scores
+of \num{0.291} and \num{0.297} are \num{-1.16} and \num{-1.10}.
+We can conclude that the \num{0.291} and \num{0.297} values are close to each other and exhibit a similar level of deviation from the mean.
 

### 2023-07-01 paper/main.tex
fix authors

@@
 \author*[1]{\fnm{Jaime} \sur{Ramírez Castillo}}\email{Jaime.Ramirez@alu.uclm.es}
+\equalcont{These authors contributed equally to this work.}
 
@@
 \equalcont{These authors contributed equally to this work.}
+
+\author[2]{\fnm{Philippe} \sur{Leray}}\email{philippe.leray@univ-nantes.fr}
+
+\affil*[1]{\orgdiv{Departamento de Sistemas Informáticos}, \orgname{Universidad de Castilla-La Mancha}, \orgaddress{\street{Campus universitario s/n}, \city{Albacete}, \postcode{02071}, \country{Spain}}}
+
+\affil[2]{\orgdiv{Laboratoire des Sciences du Numérique de Nantes}, \orgname{Centre national de la recherche scientifique, University of Nantes}, \orgaddress{\city{Nantes}, \country{France}}}
 

### 2023-07-06 paper/main.tex
improvements

@@
       \centering
-      \includegraphics[width=0.5\textwidth]{images/rmse_by_model_and_limit.png}
+      \includegraphics{images/rmse_by_model_and_limit.png}
       \caption{RMSE mean and standard deviation by model and tags/tokens limit.}
@@
       \centering
-      \includegraphics[width=0.7\textwidth]{images/rmse_by_model_and_feature.png}
+      \includegraphics{images/rmse_by_model_and_feature.png}
       \caption{RMSE mean and standard deviation by model and audio feature.}
@@
       \centering
-      \includegraphics[width=0.7\textwidth]{images/rmse_by_model_and_input.png}
+      \includegraphics{images/rmse_by_model_and_input.png}
       \caption{RMSE mean and standard deviation by model and input type (tag probablities or tokens).}
@@
       \centering
-      \includegraphics[width=0.7\textwidth]{images/rmse_by_feature_and_input.png}
+      \includegraphics[width=1\textwidth]{images/rmse_by_feature_and_input.eps}
       \caption{RMSE mean and standard deviation by audio feature and input type (tag probablities or tokens).}

### 2023-07-07 paper/main.tex
add intro remainder part

@@
 Editorial information or lyrics, for example, are mostly available in text format.
-The ability to process images or videos is also required, for example, for analyzing album artwork, or music videos.
-
-Depending on the specific MIR application, researchers or practitioners expect different output values.
-Applications that extract audio features typically return audio descriptors, namely values related to the tempo, the key, or the sample rate, among others.
+The ability to process images or videos might be also required, for example, for analyzing album artwork, or music videos.
+
+Depending on the specific MIR application, researchers and practitioners create models that generate different output values.
+Applications that extract audio features typically return audio descriptors, namely values related to the tempo, the key, or the sample rate, to name a few.
 Open source libraries such as \emph{Librosa}\footnote[1]{https://librosa.org/}
 and \emph{Essentia}\footnote[2]{https://essentia.upf.edu/index.html} offer methods to extract these values.
-Other applications might produce higher-level values, for example, by using machine learning techniques
+Other applications might produce more abstract or subjective values, for example, by using machine learning techniques
 that estimate the emotion that a track induces, or the music genre of this track.
 
-Among potentially useful input and output values, research has proved Spotify audio features and Last.fm tags to be significant values to characterize music.
-Spotify audio features capture high-level information about the music signal,
-such as energy, danceability, or valence, among others.
+Among potentially useful input and output values, research has proved \emph{Spotify audio features} and \emph{Last.fm tags} to be significant values used to  characterize music.
+Spotify audio features capture high-level information about the music signal and how humans perceive this signal.
+Examples of these features are such as energy, danceability, or valence.
 Last.fm tags are text labels that users associate to songs, artists, and albums via the Last.fm social platform.
@@
 In particular, this article focuses on predicting Spotify audio features, given a set of Last.fm tags.
-By predicting Spotify audio features, we explore the relationship between the subjective perception captured by Last.fm tags, and the concrete musical features that Spotify computes.
+By predicting Spotify audio features, we explore the relationship between the subjective text descriptions captured by Last.fm tags, and the concrete (but still user-related) musical features that Spotify computes.
 This approach might help to identify patterns and hidden correlations between how music is perceived, consumed, and discovered.
@@
 
-% The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
-% Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
-% This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.
-
-In the remainder of the article, we explain the data gathering and preparation process, as well as the data input formats and varios models.
-We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.
-\textcolor{red}{UNDER DEVELOPMENT}
-
-\section{Data Sources}
+The remainder of the article is organized as follows.
+Section \ref{sec2} covers the details Last.fm tags and Spotify Audio Features.
+Section \ref{sec3} performs a general review of previous research.
+Subsequently, section \ref{sec4} and \ref{sec5} explain how the data was gathered and prepared, as well as the conducted experiments.
+Finally, in section \ref{sec6} we conclude our study.
+% In the remainder of the article, we explain the data gathering and preparation process, as well as the data input formats and varios models.
+% We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.
+
+\section{Data Sources}\label{sec2}
 
@@
 
-\section{Related Research}
+\section{Related Research}\label{sec3}
 
@@
 
-\section{Generating a Dataset}
+\section{Generating a Dataset}\label{sec4}
 
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec5}
 
@@
       \centering
-      \includegraphics{images/rmse_by_model_and_limit.png}
+      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_limit.png}
       \caption{RMSE mean and standard deviation by model and tags/tokens limit.}
@@
       \centering
-      \includegraphics{images/rmse_by_model_and_feature.png}
+      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_feature.png}
       \caption{RMSE mean and standard deviation by model and audio feature.}
@@
       \centering
-      \includegraphics{images/rmse_by_model_and_input.png}
+      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_input.png}
       \caption{RMSE mean and standard deviation by model and input type (tag probablities or tokens).}
@@
       \centering
-      \includegraphics[width=1\textwidth]{images/rmse_by_feature_and_input.eps}
+      \includegraphics[width=1\textwidth]{images/rmse_by_feature_and_input.png}
       \caption{RMSE mean and standard deviation by audio feature and input type (tag probablities or tokens).}
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec6}
 

### 2023-07-07 paper/main.tex
improving up to 5.1

@@
 Last.fm is an online music community where users keep track of their music listening habits.
-Users apply tags to artists, tracks, and albums to categorize and describe music from their own perspective, which
-implies that the Last.fm tags space does not fit into any structured ontology or data model.
-A tag can refer to any aspect that users consider to be a valid descriptor, such as genre,
+Users apply tags to artists, tracks, and albums to categorize and describe the music they listen to, from their own perspective.
+The fact that Last.fm tags are community-contributed implies that the tag space does not fit into any structured ontology or data model.
+A tag can refer to any aspect that users consider valid and useful for the community, such as genre,
 emotion, or user listening context.
@@
 Table \ref{table:spotify-features}.
-While Spotify provides a description of the audio features.
-How they compute or estimate these values is not publicly available.
+Whereas Spotify provides a description of the audio features,
+how they compute or estimate these values is not publicly available.
 The reader can find further details about each feature in the Spotify API documentation\footnote[5]{
@@
 
-The objective of this study is to predict Spotify audio features from Last.fm tags, so the objective of this section is to
-explore existing literature that is related to the problem and review the facts that are important or relevant for our experimentation.
-
 The use of Last.fm tags and Spotify audio features has been common and prolific in studies
 that have applied machine learning to resolve MIR challenges.
-
-An important concept to consider is the arousal-valence cuadrants, present in music emotion recognition,
-related to some studies that will be reviewed in the following sections, and also related to the valence and energy
-Spotify audio features.
 
@@
 
-{\c{C}}ano and Morisio performed an analysis of Last.fm tags to create a dataset of music
+The use of Last.fm tags as input data is common in the literature.
+In particular, the Last.fm API has been tipically used to build datasets that are subsequently fed into MIR problems.
+For example, {\c{C}}ano and Morisio performed an analysis of Last.fm tags to create a dataset of music
 lyrics annotated with Last.fm tags.
@@
 In a similar direction, Bod{\'o} and Szil{\'a}gyi generated a dataset for lyrics genre classification
-by combining the Last.fm with \emph{MusicBrainz} data \cite{bodo2018connecting}.
+by combining Last.fm tags with \emph{MusicBrainz} data \cite{bodo2018connecting}.
 MusicBrainz is an online database of music editorial metadata\footnote[1]{https://musicbrainz.org/}.
 
-Although different datasets that contain Last.fm tags have been created, so far,
-the \emph{Last.fm dataset}\footnotemark[2] has been the most widely used in research.
+Among datasets that contain Last.fm tags, arguably, the \emph{Last.fm dataset}\footnotemark[2] has been the most widely used in research.
 This dataset is a complementary dataset of the Million Song Dataset (MSD) \cite{Bertin-Mahieux2011}.
@@
 
-In general, these studies confirm the possibility of extracting knowledge from Last.fm tags.
-
-\textcolor{red}{We introduced this section talking about machine learning but only refeferred to studies
-that use Last.fm tags in exploration and data set creation}
+In general, Last.fm tags have been used, generally, as input features for extracting knowledge.
+The use of Last.fm tags as target features in machine learning models, however, although less frequent, is also present \cite{eck2007autotagging}.
 
@@
 \emph{The Echo Nest} was an online music intelligence platform that provided users and clients with music analysis services.
-Among these services, the Echo Nest offered a database and an API to retrieve audio features for each of the tracks in the database \footnote[3]{https://en.wikipedia.org/wiki/The\_Echo\_Nest}.
+Among these services, the Echo Nest offered a database and an API to retrieve audio features for each of the tracks in the database\footnote[3]{https://en.wikipedia.org/wiki/The\_Echo\_Nest}.
 Spotify acquired the Echo Nest in 2014.
@@
 earlier studies use the Echo Nest denomination.
-
-Regardless of the term used, the features remain the same.
+Regardless of the term used, the list of available features remains the same.
 These features are a set of high-level descriptors, such as energy and danceability,
-which are related to the audio but also to the listeners perception.
+which are related to the audio and to the listeners perception.
 
 Similarly to Last.fm, Spotify (or Echo Nest) audio features are commonly present in MIR research.
-For example, Wang and Horv{\'a}t used  these audio features to study the differences
-between male and female artists \cite{wang2019gender}.
-\textcolor{red}{And what did they achieve?}
-
-Jamdar et al. used Echo Nest audio features, combined with lyrics data to classify songs into emotion tags.
+In one statistical study, Wang and Horv{\'a}t used these features to compare male and female artists,
+and discovered significant differences between the two binary genres \cite{wang2019gender}.
+
+Regarding its use in machine learning problems, Jamdar et al. used Echo Nest audio features, combined with lyrics data to classify songs into emotion tags.
 These classes were first defined based on a Last.fm tags emotion mapping \cite{jamdar2015emotion}.
-
-Similarly, Non-negative Matrix Factorization was applied in combination with EchoNest audio features
+In a different study, non-negative Matrix Factorization was applied in combination with EchoNest audio features
 for song recommendations \cite{benzi2016song}.
 
-Panda and Redinho explore the use of Spotify high-level features applied to Music Emotion Recognition (MER) \cite{panda2021does}.
-In particular, they identify that the energy, valence, and acousticness values, provided by the Spotify API,
-are highly relevant for emotion classification.
-They also achieve better performance on MER models by using their own top-100 features, and they determine that,
-although these three Spotify features are relevant in terms of characterizing emotion,
-more features are needed to effectively recognize emotion.
-Another interesting observation of this study is the the identification of energy as a surrogate of arousal valence
+Panda and Redinho explored the use of Spotify audio features in Music Emotion Recognition (MER) \cite{panda2021does}.
+They identified that the energy, valence, and acousticness values are highly relevant for emotion classification.
+Another interesting observation of this study is the recognition of energy as a surrogate of the valence.
+The \emph{Arousal-valence emotion plane} is an important concept in emotion-related topics, such as MER, and represents the space of emotions in a 2-dimensional plane.
 
@@
 It is unclear, however, whether these published datasets completely meet the Spotify API terms of service.
-
 For example, \emph{P4kxspotify} is a publicly available dataset that combines music review texts with Spotify audio features.
@@
 
-There are various reasons why we chose to generate our own dataset.
+\section{Generating a Dataset}\label{sec4}
+
+We have chosen to generate our own dataset for a number of reasons.
 First, we wanted to generate a dataset that combined both Last.fm tags and Spotify audio features.
 Second, there is lack of clarity regarding the conditions under which Spotify allows the use of the audio features.
-Third, we wanted to explore a research line where machine learning models are trained by using the listening history of a single user.
-% Another dataset created from the Spotify API and published to Kaggle:
-% https://www.kaggle.com/datasets/andrewmvd/spotify-playlists
-
-% TODO: move explanation to the experiment section.
-% \subsubsection{Machine Learning Models}
-
-% Several studies have explored the use of machine learning
-% models to predict audio features from audio metadata, such as Last.fm tags.
-% % which studies use Last.fm tags as parameters?
-
-% The experiments conduced in this study used two classical regression models,
-% Boosted tree regressor and Bayesian ridge regressor.
-% These two machine learning models that have been widely used in MIR. % is this true? (refs)
-
-% Boosted tree regressor is a decision tree-based model that sequentially
-% adds weak learners to the model to improve its performance.
-% Bayesian ridge regressor, on the other hand,
-% is a probabilistic model that uses Bayesian inference to estimate
-% the parameters of the model.
-
-% Additionally, language models, such as GPT-2 have shown promising results in various
-% natural language processing (NLP) and generation.
-
-% In this particular study, GPT-2 has been fine-tuned for regression tasks,
-% and has shown good performance in predicting audio features from metadata.
-
-% This paper aims to apply the boosted tree regressor,
-%  Bayesian ridge regressor, and GPT-2 models to predict Spotify audio features from Last.fm tags.
-% The experiments compare the performance of these models and evaluate their effectiveness
-% in predicting various audio features.
-% The results of these experiments will provide insights into
-% the use of different machine learning models for MIR tasks
-% and can have practical applications
-% in music recommendation systems and genre or mood recognition.
-%
-%\textcolor{red}{To be Continued \dots }
-
-\section{Generating a Dataset}\label{sec4}
+And third, we wanted to explore a research line where machine learning models are trained by using the listening history of a single user.
 
@@
 how machine learning breakthroughs have attempted, and succeeded in many cases, to generalize in a particular problem.
-This is not objective of this study, which explores how a machine learning model can represent the music consumption experience of a single human.
-The model must be able to generalize, but only within the context of the user's musical taste, which be believe can be possible, given a sufficiently large listening history.
+This is not objective of our research line, which explores how a machine learning model can represent the music consumption experience of a single human.
+Our final model must be able to generalize, but only within the context of the user's musical taste, which be believe can be possible, given a sufficiently large listening history.
 
@@
 Last.fm started monitoring user listening activity with a desktop application called \emph{Scrobbler}.
-Users used to install this application on their computers to monitor their activity on players such as Winamp or iTunes.
+Users install this application on their computers to monitor their activity on players such as Winamp or iTunes.
 With the advent of music streaming services, the possibilities for users to scrobble their music habits expanded.
@@
 We queried the Last.fm API to download the user{'}s scrobbling logs, reported from 2007 to 2022.
-For each scrobble, we have gathered the following information:
+For each scrobble, we gathered the following information:
 
@@
 After collecting the listening history and track tags from Last.fm, and identifying the unique
-tracks that represent the user music collection, we used the Spotify API to
+tracks that represent the user music collection, we queried the Spotify API to
 collect audio features for each one of the \num{20000} individual tracks.
@@
 For each track, the artist and track name extracted from Last.fm were used as parameters of the Spotify Search API.
-The Last.fm API provides a unique identifier, the \emph{MusicBrainz} ID.
-The Spotify API, however, does not provide this value.
+The Last.fm API provides a unique identifier, the \emph{MusicBrainz} ID for some tracks.
+The Spotify API, however, does not provide this value so we could not establish an unequivocal mapping.
 
@@
 After retrieving audio features, we identified that the Spotify API had failed to provide audio features for a portion of the tracks.
-Similarly, Last.fm also returned no tags for another subset of the tracks.
+Similarly, Last.fm returned an empty tag list for another subset of the tracks.
 To prevent problems with missing values, we decided to filter out these tracks from the dataset.
@@
 This model has proved to be a powerful solution for classification and regression problems on high-dimensional, structured data.
-\textcolor{red}{REF}
 
